### 2021-05-12 IEEE SP Magazine/05-12-2021 - Draft/article.tex
Updated the article.

\documentclass[journal]{IEEEtran}

\usepackage{amsmath,cite,graphicx}

\begin{document}

\title{The Constant-Q Transform \\ Spectral Envelope Coefficients}

\maketitle

\section{Scope}

\IEEEPARstart{T}{he} sliding discrete Fourier transform (SDFT) ...

%One paragraph describing the article content

\cite{brown1991}, \cite{brown1992}

\cite{engel2017}: NSynth

- Aware that some methods are attempting to learn more adapted feature like that 
- The envelope component can additionally be refined, along with the pitch component
- Incidentally, the pitch component can be used to identify the pitch/key

\section{Relevance}

%one or two paragraphs

\section{Prerequisites}

Basic knowledge of audio signal processing and music information retrieval is required to understand this article, in particular, concepts such as the Fourier transform, convolution, spectral envelope, pitch, CQT, and MFCC.

\section{Problem Statement}

\subsection{Observations}

Assumption: A log-spectrum, such as the CQT-spectrum, can be represented as the convolution of a pitch-invariant log-specrtal envelope component (= timbre) and a envelope-normalized pitch component.

\begin{itemize}
\item A pitch change in the audio translates to a linear shift in the log-spectrum.
\item The Fourier transform (FT) of a convolution of two functions is equal to the point-wise product of their FTs (convolution theorem).
\item The magnitude FT is shift-invariant.
\end{itemize}

%Equation \ref{eq:sdft} shows ...
%
%\begin{equation}
%\label{eq:sdft}
%\begin{split}
%\underset{0 \leq k < N}{X_k^{(i)}} 
%& = \sum_{n=0}^{N-1} x_{i+n} e^{\frac{-j 2\pi n k}{N}}\\
%& = \sum_{n=0}^{N-1} x_{i+n} e^{\frac{-j 2\pi (n+1) k}{N}} e^{\frac{j 2\pi k}{N}} 
%\end{split}
%\end{equation}

\section{Solution}

%\begin{figure}[htbp]
%	\centering
%	\includegraphics[width=\columnwidth]{dft_kernels.png}
%	\caption{Kernels derived from the Hanning (top-left), Blackman (top-right), triangular (center-left), Parzen (center-right), Gaussian (with $\alpha = 2.5$) (bottom-left), and Kaiser (with $\beta = 0.5$) (bottom-right) windows. The kernels were derived for an $N$-point DFT where $N = 2048$ samples. Only the first 100 coefficients at the bottom-left corner of the $N$-by-$N$ kernels are shown. The values are displayed in log of amplitude.}
%	\label{fig:dft_kernels}
%\end{figure}

%Figure \ref{fig:dft_kernels} shows ...

\section{Numerical Example}

% or "V - Computational Example"

\section{What We Have Learned}

We have shown that ...

% One paragraph

\section{Author}

\textit{\textbf{Zafar Rafii}} (zafarrafii@gmail.com) received a PhD in Electrical Engineering and Computer Science from Northwestern University in 2014, and an MS in Electrical Engineering from both Ecole Nationale Superieure de l’Electronique et de ses Applications in France and Illinois Institute of Technology in the US in 2006. He is currently a senior research engineer at Gracenote in the US. He also worked as a research engineer at Audionamix in France. His research interests are centered on audio analysis, somewhere between signal processing, machine learning, and cognitive science, with a predilection for source separation and audio identification.

\bibliographystyle{IEEEtran}
\bibliography{bibliography.bib}

\end{document}

### 2021-05-13 IEEE SP Magazine/05-2021 - Draft/article.tex
Started writing intro in draft.

--- 2021-05-12 IEEE SP Magazine/05-12-2021 - Draft/article.tex
@@
 
-\title{The Constant-Q Transform \\ Spectral Envelope Coefficients}
+\title{The Constant-Q Transform Spectral Envelope Coefficients: a musically-adapted timbre feature}
 
@@
 \section{Scope}
-
-\IEEEPARstart{T}{he} sliding discrete Fourier transform (SDFT) ...
-
 %One paragraph describing the article content
 
+%- Aware that some methods are attempting to learn more adapted feature like that 
+%- The envelope component can additionally be refined, along with the pitch component
+%- Incidentally, the pitch component can be used to identify the pitch/key
+
+% We are presenting a timbre feature better suited to musical data
+% We will show how to derive it from the CQT, by decomposing it into a pitch-invariant spectral envelop and a normalized pitch component
+% We will show how it compares to the MFCC, commonly-used as the timbre feature, on the NSynth dataset
+
+\IEEEPARstart{T}{imbre} is the attribute of sound which makes, for example, two musical instruments playing the same note sound different.
+It is generally associated with the spectral (but also temporal) envelope and is typically assumed to be independent from the pitch (but also the loudness) of the sound \cite{moore2004}.
+This article will show how to derive a simple but ...
+timbre feature ...
+
 \cite{brown1991}, \cite{brown1992}
-
 \cite{engel2017}: NSynth
 
-- Aware that some methods are attempting to learn more adapted feature like that 
-- The envelope component can additionally be refined, along with the pitch component
-- Incidentally, the pitch component can be used to identify the pitch/key
+\section{Relevance}
+%One or two paragraphs
 
-\section{Relevance}
-
-%one or two paragraphs
+% A good pitch-invariant timbre feature can help for identifying musical instruments 
+% Additionally, we will show how to decompose the CQT spectrum into a pitch-invariant envelope and a normalized pitch component, which can be reused
 
@@
 \section{Problem Statement}
+
+% There is no proper timbre feature, especially for music data.
+% People use MFCC, originally designed for speech data, other tried to learn such features
+%
+
+\section{Solution}
 
@@
 \end{itemize}
+
+\section{Numerical Example}
+% or "V - Computational Example"
 
@@
 
-\section{Solution}
+%Figure \ref{fig:dft_kernels} shows ...
 
@@
 
-%Figure \ref{fig:dft_kernels} shows ...
-
-\section{Numerical Example}
-
-% or "V - Computational Example"
-
 \section{What We Have Learned}
+% One paragraph
 
 We have shown that ...
-
-% One paragraph
 

### 2021-05-17 IEEE SP Magazine/05-2021 - Draft/article.tex
Worked a bit more on the intro.

@@
 It is generally associated with the spectral (but also temporal) envelope and is typically assumed to be independent from the pitch (but also the loudness) of the sound \cite{moore2004}.
-This article will show how to derive a simple but ...
-timbre feature ...
+While a complex concept to encompass, this article will show how to compute a simple but representative timbre feature ...
+well-adapted to musical data.
 
-\cite{brown1991}, \cite{brown1992}
-\cite{engel2017}: NSynth
+The feature will be derived from the constant-Q transform (CQT), a log-scaled frequency transform better adapted to musical data \cite{brown1991}, \cite{brown1992}.
+
+The CQT-SEC will compare with the mel-frequency cepstral coefficients (MFCC), on the NSynth dataset, a publicly-available dataset of musical notes \cite{engel2017}.
 

### 2021-05-18 IEEE SP Magazine/05-2021 - Draft/SDFT.tex
Still working on the intro.

\documentclass[journal]{IEEEtran}

\usepackage{amsmath,cite,graphicx}

\begin{document}

\title{Sliding DFT with Kernel Windowing}

\maketitle

\section{Scope}

\IEEEPARstart{T}{he} sliding discrete Fourier transform (SDFT) is an efficient method for computing the $N$-point discrete Fourier transform (DFT) of a given signal starting at a given sample from the $N$-point DFT of the same signal starting at the previous sample \cite{jacobsen2003}. However, the SDFT does not allow the use of a window function, generally incorporated in the computation of the DFT to reduce spectral leakage, as it would break its sliding property. This lecture-notes article will show how windowing can be included in the SDFT by using a kernel derived from the window function, while keeping the process computationally efficient. In addition, this approach allows to turn other transforms, such as the modified discrete cosine transform (MDCT), into efficient sliding versions of themselves.

\section{Relevance}

The SDFT can be used to perform spectral analysis on successive samples in a signal without having to compute a new DFT from scratch every time, provided that windowing can be incorporated into the computation of the DFTs without harming the efficiency of the method. A notable application of the SDFT with windowing can then be framing detection in audio signals which have undergone lossy compression, in the context of audio compression identification \cite{kim2018}. A lossy compression algorithm will typically introduce traces of compression in the signal being encoded, which can become visible in the time-frequency representation when using the same parameters and framing that were used for the encoding. The parameters and framing can therefore be recovered by computing time-frequency representations at successive samples in the signal and identifying when traces of compression become visible. This demanding process can be translated into an efficient one by using the SDFT with kernel windowing.

\section{Prerequisites}

Basic knowledge of digital signal processing is required to understand this article, in particular, concepts such as the DFT, windowing, and general spectral analysis. More details about the SDFT and lossy audio compression identification can also be found in \cite{jacobsen2003} and \cite{kim2018}, respectively.

\section{Problem Statement and Solution}

\subsection{Problem Statement}

The SDFT allows to compute the $N$-point DFT of a signal from the $N$-point DFT of the same signal starting one sample earlier, in a sense by sliding a rectangular window of length $N$ one sample forward. The SDFT essentially relies on the shift theorem which states that multiplying a signal by a linear phase is equivalent to a circular shift in the corresponding DFT.

Equation \ref{eq:sdft} shows the derivation of $X^{(i)}$, the $N$-point DFT of signal $x$ starting at sample $i$, from $X^{(i-1)}$, the $N$-point DFT of $x$ starting at $i-1$, process hence known as SDFT.
\begin{equation}
\label{eq:sdft}
\begin{split}
\underset{0 \leq k < N}{X_k^{(i)}} 
& = \sum_{n=0}^{N-1} x_{i+n} e^{\frac{-j 2\pi n k}{N}}\\
& = \sum_{n=0}^{N-1} x_{i+n} e^{\frac{-j 2\pi (n+1) k}{N}} e^{\frac{j 2\pi k}{N}} \\
& = \sum_{n=1}^{N} x_{i-1+n} e^{\frac{-j 2\pi n k}{N}} e^{\frac{j 2\pi k}{N}} \\
& = \left( \sum_{n=0}^{N-1} x_{i-1+n} e^{\frac{-j 2\pi n k}{N}} - x_{i-1} + x_{i+N-1} \right) e^{\frac{j 2\pi k}{N}} \\
& = \left(X_k^{(i-1)} - x_{i-1} + x_{i+N-1} \right) e^{\frac{j 2\pi k}{N}}
\end{split}
\end{equation}

The SDFT thus only requires $2N$ additions and $N$ multiplications, leading to a linear time complexity of $\mathcal{O}(N)$, while the full and direct computation of the DFT and the fast Fourier transform (FFT) are $\mathcal{O}(N^2)$ and $\mathcal{O}(N \log N)$, respectively.

Transforms such as the DFT typically use a window function in their computation to reduce spectral leakage and enhance spectral analysis. However, the SDFT does not allow the incorporation of a window function as it will break the process shown in Equation \ref{eq:sdft}. One solution would be to perform the windowing in the frequency domain, that is, on the derived DFT through convolution. A practical window function for that matter could be the Hanning window, as the corresponding windowing in the frequency domain equals a simple three-point convolution \cite{jacobsen2003}. Other window functions however may not be as practical, as the corresponding convolutions may involve many more operations, which will ultimately hurt the computational efficiency of the SDFT. The problem is therefore to incorporate any window function into the computation of the DFTs in an efficient manner without breaking the SDFT process.

\subsection{Solution: Kernel Windowing}

The idea of performing the windowing in the frequency domain can still be exploited, by reformulating the convolution as a multiplication by a kernel which can be derived from any window function. Such a kernel will be independent from the signal to be processed and only need to be computed once. It will typically have a very small number of values which would be significant, which means that most of the values can then be ignored. This would lead to a very sparse kernel which can then be applied to the DFT of the signal, producing results virtually equivalent to the DFT of the same signal modified by the corresponding window function, while preserving the computational efficiency of the SDFT.

The constant-Q transform (CQT) is a transform with a logarithmic frequency resolution which was proposed as a more adapted alternative to the Fourier transform for analyzing music signals \cite{brown1991}. A fast algorithm was soon after proposed which translated the slow computation of the CQT into the multiplication of a DFT and a kernel, the former one which can be efficiently computed using the FFT and the latter one which is computed once beforehand and is typically very sparse \cite{brown1992}. The idea was to use Parseval's theorem to turn the direct computation in the time domain into a multiplication between a DFT and a kernel in the frequency domain, essentially demonstrating the property of energy conversation between the time and the frequency domains \cite{oppenheim1975}.

Equation \ref{eq:parseval} recalls Parseval's theorem. $X$ is the $N$-point DFT of $x$ and $\overline{x}$ represents the complex conjugate of $x$.
\begin{equation}
\label{eq:parseval}
\sum_{n=0}^{N-1} x_n \overline{y_n} = \frac{1}{N} \sum_{k=0}^{N-1} X_k \overline{Y_k}
\end{equation}

Following a similar idea, we propose to use Parseval's theorem to translate the DFT of a windowed signal into the DFT of the signal, multiplied by a kernel which is derived from the corresponding window function, a multiplication which will happen after the SDFT process. Unlike in the fast CQT case, the purpose here is not to speed up the computation of the transform by taking advantage of the efficiency of the FFT algorithm in conjunction with the use of sparse kernel, but to extract the windowing operation from the DFT computation, so that the SDFT process shown in Equation \ref{eq:sdft} still holds.

Equation \ref{eq:sliding_dft_windowing} shows the computation of $\mathcal{X}^{(i)}$, the $N$-point DFT of the signal $x$ starting at sample $i$ and modified by the window function $w$, from $X^{(i-1)}$, the $N$-point DFT of $x$ starting at $i-1$ (without windowing), multiplied by the kernel $K$ which is derived from $w$.

\begin{equation}
\label{eq:sliding_dft_windowing}
\begin{split}
\underset{0 \leq k < N}{{\mathcal{X}_k}^{(i)}}
& = \sum_{n=0}^{N-1} x_{i+n} \underbrace{w_n e^{\frac{-j 2\pi n k}{N}}}_{\overline{y_n}} \\
& = \sum_{k'=0}^{N-1} X_{k'}^{(i)} \underbrace{K_{k,k'}}_{\frac{1}{N} \overline{Y_{k'}}} \\
& = \sum_{k'=0}^{N-1} \left[\left(X_{k'}^{(i-1)} - x_{i-1} + x_{i+N-1} \right) e^{\frac{j 2\pi k'}{N}}\right] K_{k,k'} \\
\underset{0 \leq k' < N}{\underset{0 \leq k < N}{K_{k,k'}}} 
& = \frac{1}{N} \overline{Y_{k'}} = \frac{1}{N} \overline{\sum_{n=0}^{N-1} y_n  e^{\frac{-j 2\pi n k'}{N}}} \\
& = \frac{1}{N} \overline{\sum_{n=0}^{N-1} \overline{w_n e^{\frac{-j 2\pi n k}{N}}}  e^{\frac{-j 2\pi n k'}{N}}} \\
& = \frac{1}{N} \sum_{n=0}^{N-1} w_n e^{\frac{j 2\pi n (k'-k)}{N}}
\end{split}
\end{equation}

As we can see in Equation \ref{eq:sliding_dft_windowing}, the kernel is completely independent from the signal, therefore it only needs to be computed once, before the SDFT process. Furthermore, given the nature of such kernel, typically only a very small number of its values will be significant, which means that most of the values can then be zeroed (given some threshold) leading to a very sparse kernel. The multiplication of the derived DFT by such kernel will thus only involve few more operations, keeping the whole process computationally efficient.

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\columnwidth]{dft_kernels.png}
	\caption{Kernels derived from the Hanning (top-left), Blackman (top-right), triangular (center-left), Parzen (center-right), Gaussian (with $\alpha = 2.5$) (bottom-left), and Kaiser (with $\beta = 0.5$) (bottom-right) windows. The kernels were derived for an $N$-point DFT where $N = 2048$ samples. Only the first 100 coefficients at the bottom-left corner of the $N$-by-$N$ kernels are shown. The values are displayed in log of amplitude.}
	\label{fig:dft_kernels}
\end{figure}

Figure \ref{fig:dft_kernels} shows the kernels derived from some common window functions, namely Hanning, Blackman, triangular, Gaussian, Parzen, and Kaiser windows. As we can see, the Hanning window kernel shows only three non-zero values per row, confirming that the corresponding windowing in the frequency domain equals a simple three-point convolution, while the Blackman window kernel shows five non-zero values per row. Both those windows are actually special cases of the generalized cosine window whose corresponding windowing in the frequency domain equals convolutions with typically only few points. Unlike the Hanning and Blackman window kernels, the triangular, Parzen, Gaussian, and Kaiser window kernels show additional non-zero values around their main diagonal, suggesting that the corresponding windowings in the frequency domain equal convolutions with many more points. However, most of those non-zero values have very small magnitudes ($\ll 0.01$) and could then be ignored without significantly affecting the actual windowing process. By using an appropriate threshold, those kernels can therefore be made very sparse with only a few meaningful values per row, in the same manner as in the fast CQT case \cite{brown1992}.

As proposed in \cite{brown1992}, we computed for each of those kernels the error in keeping the values greater than a chosen threshold, by dividing the sum of the magnitudes of the values after thresholding by the sum of the magnitudes of all the values before thresholding. A threshold of $0.01$ will thus give very small errors of $0.049$, $0.009$, $0.020$, and $0.015$, for the triangular, Parzen, Gaussian, and Kaiser window kernels, respectively, when derived for an $N$-point DFT with $N = 2048$. With such threshold, the first three kernels will then only have five non-zero values per row, while the latter one will have three non-zero values per row. This shows that only a very small number of values are actually significant in such kernels. The multiplication of the DFT by those very sparse kernels will then only involve $KN$ multiplications and $KN$ additions, with $K=3$ or $5$, barely affecting the computational efficiency of the SDFT, still maintaining a linear complexity of $\mathcal{O}(N)$, and producing results virtually equivalent to taking the DFT of the signal modified by the corresponding window functions.

\section{Computational Examples}

\subsection{Framing Detection and Lossy Audio Coding}

The SDFT with kernel windowing can be particularly useful for fast framing detection, in the context of audio compression identification. Audio compression identification is the recovery of information regarding the data compression that an audio signal has undergone. In particular, the recovery of the parameters and framing used at the time-frequency decomposition stage of the encoding could allow to identify the coding format or detect alterations in audio signals which have undergone lossy compression \cite{herre2000,moehrs2002,yang2008,gartner2014,kim2018}. Lossy compression algorithms typically introduce traces of compression in the audio signal being encoded, in the form of time-frequency coefficients quantized to zeros which can become visible when using the same parameters and framing that were used for the encoding. One approach to identify if and when lossy compression was used would then be to compute the time-frequency representation at successive samples in the audio signal and search for traces of compression every time, given a set of parameters associated with a known coding format, such as time-frequency transform, window length, and window function, a process also known as framing detection.

Lossy audio coding formats, perhaps the most popular ones being MP3, AAC, AC-3, Vorbis, and WMA, are widely used for storage, e.g., in music and video files, or transmission, e.g., in radio and television broadcasting. Compression algorithms which can encode to such formats first transform the audio signal into a time-frequency representation, derive a psychoacoustic model to locate regions of perceptually less significance, then quantize the data given the psychoacoustic model, and finally convert it into a bitstream. The transform used at the time-frequency decomposition stage is typically based on the MDCT, and a variety of window lengths and window functions can be used depending on the coding format. In particular, specialized window functions such as the sine, slope and Kaiser-Bessel-derived (KBD) windows are generally required for the MDCT to be invertible. For more information about lossy audio coding, the reader is referred to \cite{bosi2003}.

Equation \ref{eq:mdct} shows the computation of the MDCT (without windowing).

\begin{equation}
\label{eq:mdct}
\underset{0 \leq k < \frac{N}{2}}{Y_k}  = \sum_{n=0}^{N-1} x_n cos \left[\frac{2\pi}{N} \left(n+\frac{1}{2}+\frac{N}{4}\right)\left(k+\frac{1}{2}\right)\right] \\
\end{equation}

\subsection{Sliding MDCT with kernel windowing}

In this context, performing framing detection for audio compression identification would involve computing an MDCT for every set of window length and window function associated with a known lossy coding format, at successive samples in the audio signal, and search for time-frequency coefficients quantized to zero, until one of the set shows visible traces of compression for a specific framing of the signal. We can see that such process will be computationally demanding, as a full transform would have to be computed every time. The direct computation of the MDCT, including the windowing using one of the specialized window functions presented earlier, can actually be translated into an SDFT with kernel windowing, by incorporating the computation of the window function, and a part of the MDCT, into a kernel which will still happen to be very sparse, thus making the process computationally efficient.

Equation \ref{eq:sliding_mdct} shows the computation of $\mathcal{Y}^{(i)}$, the $N$-point MDCT of signal $x$ starting at sample $i$ and modified by the window function $w$, from $X^{(i-1)}$, the $N$-point DFT of $x$ starting at $i-1$ (without windowing), multiplied by the kernel $K$ which is derived from $w$.
\begin{equation}
\label{eq:sliding_mdct}
\begin{split}
\underset{0 \leq k < \frac{N}{2}}{\mathcal{Y}_k^{(i)}} 
& = \sum_{n=0}^{N-1} x_{i+n} \underbrace{w_n cos \left[\frac{2\pi}{N} \left(n+\frac{1}{2}+\frac{N}{4}\right)\left(k+\frac{1}{2}\right)\right]}_{\overline{y_n}} \\
& = \sum_{k'=0}^{N-1} X_{k'}^{(i)} \underbrace{K_{k,k'}}_{\frac{1}{N} \overline{Y_{k'}}} \\
& = \sum_{k'=0}^{N-1} \left[\left(X_{k'}^{(i-1)} - x_{i-1} + x_{i+N-1} \right) e^{\frac{j 2\pi k'}{N}}\right] K_{k,k'} \\
\underset{0 \leq k' < N}{\underset{0 \leq k < \frac{N}{2}}{K_{k,k'}}} 
& = \frac{1}{N} \overline{Y_{k'}} = \frac{1}{N} \overline{\sum_{n=0}^{N-1} y_n  e^{\frac{-j 2\pi n k'}{N}}} \\
& = \frac{1}{N} \overline{\sum_{n=0}^{N-1} \overline{w_n cos  \left[\!\frac{2\pi}{N}\!\left(\!n\!+\frac{1}{2}\!+\!\frac{N}{4}\!\right)\!\left(\!k\!+\!\frac{1}{2}\!\right)\!\right]} e^{\frac{-j 2\pi n k'}{N}}} \\
& = \frac{1}{N} \sum_{n=0}^{N-1} w_n cos  \left[\!\frac{2\pi}{N}\!\left(\!n\!+\frac{1}{2}\!+\!\frac{N}{4}\!\right)\!\left(\!k\!+\!\frac{1}{2}\!\right)\!\right]  e^{\frac{j 2\pi n k'}{N}} \\
\end{split}
\end{equation}

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\columnwidth]{mdct_kernels.png}
	\caption{Kernels derived for an $N$-point MDCT, from the sine window where $N = 1152$ samples (top), the slope window where $N = 2048$ samples (center), and the KBD window where $N = 512$ samples (bottom). The values are displayed in log of amplitude.}
	\label{fig:mdct_kernels}
\end{figure}

Figure \ref{fig:mdct_kernels} shows the kernels derived for an N-point MDCT, from the sine window where $N = 1152$ samples, as in MP3, from the slope window where $N = 2048$ samples, as in Vorbis, and from the KBD window where $N = 512$ samples, as in AC-3. As we can see, most of the values in those kernels appear to have negligible magnitudes, while the very few values with significant magnitudes appear to be concentrated around two diagonals, one going from the bottom-left to the top-center and one going from the top-center to the bottom-right. As in \cite{brown1992}, we computed for each of those kernels the error in keeping the values greater than $0.01$ and obtained very small errors of $0.000$, $0.022$, and $0.013$, for the sine, slope, and KBD window kernel, respectively. With such threshold, the sine kernel will only have around two non-zero values per row, and the slope and KBD kernels around six non-zero values per row. These very sparse kernels will therefore barely affect the computational efficiency of the SDFT, while still producing results equivalent to taking the MDCT of the signal modified by the corresponding window functions.

\section{What We Have Learned}

We have shown that the SDFT can incorporate windowing in its computation by using a kernel which can be derived from any window function and which can be made very sparse. This SDFT with kernel windowing will produce results equivalent to the DFT of the signal modified by the corresponding window function, while keeping the process computationally efficient. This approach can find application in audio compression identification, in particular by making the process of framing detection much more efficient, allowing to translate a transform such as the MDCT into an efficient sliding version of itself.

\section{Author}

\textit{\textbf{Zafar Rafii}} (zafar.rafii@nielsen.com) is a senior research engineer at Gracenote, with a PhD in Electrical Engineering and Computer Science from Northwestern University.

\bibliographystyle{IEEEtran}
\bibliography{bibliography}

\end{document}

### 2021-05-18 IEEE SP Magazine/05-2021 - Draft/article.tex
Still working on the intro.

--- 2021-05-13 IEEE SP Magazine/05-2021 - Draft/article.tex
@@
 
-\title{The Constant-Q Transform Spectral Envelope Coefficients: a musically-adapted timbre feature}
+\title{The Constant-Q Transform Spectral Envelope Coefficients: A Timbre Feature Designed for Music}
 
@@
 \section{Scope}
-%One paragraph describing the article content
-
-%- Aware that some methods are attempting to learn more adapted feature like that 
-%- The envelope component can additionally be refined, along with the pitch component
-%- Incidentally, the pitch component can be used to identify the pitch/key
-
-% We are presenting a timbre feature better suited to musical data
-% We will show how to derive it from the CQT, by decomposing it into a pitch-invariant spectral envelop and a normalized pitch component
-% We will show how it compares to the MFCC, commonly-used as the timbre feature, on the NSynth dataset
 
@@
 It is generally associated with the spectral (but also temporal) envelope and is typically assumed to be independent from the pitch (but also the loudness) of the sound \cite{moore2004}.
-This article will show how to derive a simple but ...
-timbre feature ...
+Typical attempts to characterize the timbre of musical data is to use to the mel-frequency cepstral coefficients (MFCC) \cite{davis1980}, a feature original designed for speech recognition, while more recently, data-driven approach ...
 
-\cite{brown1991}, \cite{brown1992}
-\cite{engel2017}: NSynth
+In this article, we will show how to compute a timbre feature that is well-adapted to musical data.
+The feature will be derived from the constant-Q transform (CQT), a log-scaled frequency transform which matches the notes of the Western music scale \cite{brown1991}, \cite{brown1992}.
+
+Decompose the CQT into a pitch-invariant spectral envelope and an energy-normalized pitch component.
+
+CQT spectral envelope coefficients (CQT-SEC)
+
+The CQT-SEC will compare with the , on the NSynth dataset, a publicly-available dataset of musical notes \cite{engel2017}.
+
+%interpretable feature
 
@@
 %One or two paragraphs
+
+% A good timbre feature can be used 
+% - can be used for similarity, classification, as inputs to supervised classifiers, as a preprocessing step, ...
 
@@
 
+%- The envelope component can additionally be refined, along with the pitch component
+%- Incidentally, the pitch component can be used to identify the pitch/key
+
+convolution theorem: \cite{proakis1995}.
+
 \subsection{Observations}
@@
 \end{itemize}
+
+% Can think about the extraction of the envelop as a white normalization of the pitch by using a weighting function that normalizes the signal spectral density by the spectrum magnitude, just like in the GCC-PHAT
 

### 2021-05-18 IEEE SP Magazine/05-2021 - Draft/article.tex
Almost done with the intro.

@@
 It is generally associated with the spectral (but also temporal) envelope and is typically assumed to be independent from the pitch (but also the loudness) of the sound \cite{moore2004}.
-Typical attempts to characterize the timbre of musical data is to use to the mel-frequency cepstral coefficients (MFCC) \cite{davis1980}, a feature original designed for speech recognition, while more recently, data-driven approach ...
-
-In this article, we will show how to compute a timbre feature that is well-adapted to musical data.
-The feature will be derived from the constant-Q transform (CQT), a log-scaled frequency transform which matches the notes of the Western music scale \cite{brown1991}, \cite{brown1992}.
-
-Decompose the CQT into a pitch-invariant spectral envelope and an energy-normalized pitch component.
-
-CQT spectral envelope coefficients (CQT-SEC)
-
-The CQT-SEC will compare with the , on the NSynth dataset, a publicly-available dataset of musical notes \cite{engel2017}.
-
-%interpretable feature
+In this article, we will show how to derive a simple but well-founded timbre feature from the constant-Q transform (CQT), a log-scaled frequency transform that is well-adapted to musical data \cite{brown1991}, \cite{brown1992}.
+We will show how to decompose the CQT into an energy-normalized pitch component and a pitch-invariant spectral envelope, from which we will extract a number of meaningful coefficients.
+We will then compare these CQT spectral envelope coefficients (CQT-SEC) with the mel-frequency cepstral coefficients (MFCC) \cite{davis1980}, a feature originally designed for speech recognition but liberally used to characterize timbre in music, on the NSynth dataset, a large-scale and publicly-available dataset of musical notes \cite{engel2017}.
 

### 2021-05-19 IEEE SP Magazine/05-2021 - Draft/article.tex
Finished scope, relevance, and prerequisites; working on problem statement.

@@
 It is generally associated with the spectral (but also temporal) envelope and is typically assumed to be independent from the pitch (but also the loudness) of the sound \cite{moore2004}.
-In this article, we will show how to derive a simple but well-founded timbre feature from the constant-Q transform (CQT), a log-scaled frequency transform that is well-adapted to musical data \cite{brown1991}, \cite{brown1992}.
-We will show how to decompose the CQT into an energy-normalized pitch component and a pitch-invariant spectral envelope, from which we will extract a number of meaningful coefficients.
-We will then compare these CQT spectral envelope coefficients (CQT-SEC) with the mel-frequency cepstral coefficients (MFCC) \cite{davis1980}, a feature originally designed for speech recognition but liberally used to characterize timbre in music, on the NSynth dataset, a large-scale and publicly-available dataset of musical notes \cite{engel2017}.
+In this article, we will show how to design a simple but efficient pitch-independent timbre feature which is well-adapted to musical data, by deriving it from the constant-Q transform (CQT) \cite{brown1991}, \cite{brown1992}, a log-scaled frequency transform which matches the notes of the Western music scale.
+We will show how to decompose the CQT into an energy-normalized pitch component and a pitch-invariant spectral envelope, the latter from which we will extract a number of meaningful coefficients.
+We will then evaluate the discriminative power of these CQT spectral envelope coefficients (CQT-SEC) on the NSynth dataset \cite{engel2017}, a large-scale dataset of musical notes which is publicly available, comparing them with the mel-frequency cepstral coefficients (MFCCs) \cite{davis1980}, features originally designed for speech recognition but commonly used to characterize timbre in music.
 
 \section{Relevance}
-%One or two paragraphs
 
-% A good timbre feature can be used 
-% - can be used for similarity, classification, as inputs to supervised classifiers, as a preprocessing step, ...
-
-% A good pitch-invariant timbre feature can help for identifying musical instruments 
-% Additionally, we will show how to decompose the CQT spectrum into a pitch-invariant envelope and a normalized pitch component, which can be reused
+A timbre feature which is well-adapted to musical data, pitch-independent, and with high discriminative power can be helpful for a number of applications, such as computing similarity, allowing identification, and performing classification of musical data, in particular, musical instruments.
+Additionally, the ability to decompose the spectrum of a sound (here, the magnitude CQT) into a pitch-independent spectral envelope and an energy-normalized pitch component, can be useful for music analysis, transformation, and resynthesis.
+The energy-normalized pitch component can also potentially be used for pitch identification and melody extraction.
 
@@
 
-Basic knowledge of audio signal processing and music information retrieval is required to understand this article, in particular, concepts such as the Fourier transform, convolution, spectral envelope, pitch, CQT, and MFCC.
+Basic knowledge of audio signal processing and music information retrieval (MIR) is required to understand this article, in particular, concepts such as the Fourier transform, convolution, spectral envelope, pitch, CQT, and MFCCs.
 
@@
 
-% There is no proper timbre feature, especially for music data.
-% People use MFCC, originally designed for speech data, other tried to learn such features
-%
+The complex nature of timbre makes it an attribute that is not easy to quantify in terms of one single characteristic feature.
+While it is assumed to be independent from pitch and loudness, it is not really possible to fully disentangle timbre from those qualities, as timbre is inherently dependent on the spectral content of the sound, which is also defined by its pitch and loudness \cite{moore2004}.
+Researchers in MIR proposed a number of descriptors to characterize one or more aspects of timbre, but ...
+mostly adopted the MFCCs
+
+%Given a lack of proper timbre feature for characterizing musical data, the MIR community adopted the MFCCs,
+
+% MIR tried some feature linked to timbre but nothing that emcompasses the whole attribute.
+Adopted MFCCs from speech recognition, commonly-used.
+While more recent data-driven approaches are attempting to learn a timbre feature
 
@@
 
+%Limitations:
+%- just like MFCCs, not very robust to noise, to mixture of musical instruments playing at different pitches
+
 \section{Author}

### 2021-05-19 IEEE SP Magazine/05-2021 - Draft/article.tex
Revisions.

@@
 It is generally associated with the spectral (but also temporal) envelope and is typically assumed to be independent from the pitch (but also the loudness) of the sound \cite{moore2004}.
-In this article, we will show how to design a simple but efficient pitch-independent timbre feature which is well-adapted to musical data, by deriving it from the constant-Q transform (CQT) \cite{brown1991}, \cite{brown1992}, a log-scaled frequency transform which matches the notes of the Western music scale.
-We will show how to decompose the CQT into an energy-normalized pitch component and a pitch-invariant spectral envelope, the latter from which we will extract a number of meaningful coefficients.
+In this article, we will show how to design a pitch-independent timbre feature which is well-adapted to musical data, by deriving it from the constant-Q transform (CQT) \cite{brown1991}, \cite{brown1992}, a log-scaled frequency transform which matches the notes of the Western music scale.
+We will show how to decompose the magnitude CQT spectrum into an energy-normalized pitch component and a pitch-independent spectral envelope, the latter from which we will extract a number of timbral coefficients.
 We will then evaluate the discriminative power of these CQT spectral envelope coefficients (CQT-SEC) on the NSynth dataset \cite{engel2017}, a large-scale dataset of musical notes which is publicly available, comparing them with the mel-frequency cepstral coefficients (MFCCs) \cite{davis1980}, features originally designed for speech recognition but commonly used to characterize timbre in music.
@@
 
-A timbre feature which is well-adapted to musical data, pitch-independent, and with high discriminative power can be helpful for a number of applications, such as computing similarity, allowing identification, and performing classification of musical data, in particular, musical instruments.
-Additionally, the ability to decompose the spectrum of a sound (here, the magnitude CQT) into a pitch-independent spectral envelope and an energy-normalized pitch component, can be useful for music analysis, transformation, and resynthesis.
+A timbre feature which is well-adapted to musical data, pitch-independent, and with high discriminative power can find uses in a number of applications, such as similarity detection, sound recognition, and music classification, in particular, of musical instruments.
+Additionally, the ability to decompose the spectrum of a sound (here, the magnitude CQT) into a pitch-independent spectral envelope and an energy-normalized pitch component can be useful for audio analysis, transformation, and resynthesis.
 The energy-normalized pitch component can also potentially be used for pitch identification and melody extraction.
@@
 
-Basic knowledge of audio signal processing and music information retrieval (MIR) is required to understand this article, in particular, concepts such as the Fourier transform, convolution, spectral envelope, pitch, CQT, and MFCCs.
+Basic knowledge of audio signal processing and some knowledge of music information retrieval (MIR) are required to understand this article, in particular, concepts such as the Fourier transform, convolution, spectral envelope, pitch, CQT, and MFCCs.
 

### 2021-05-20 IEEE SP Magazine/05-2021 - Draft/article.tex
Working on the problem section.

@@
 It is generally associated with the spectral (but also temporal) envelope and is typically assumed to be independent from the pitch (but also the loudness) of the sound \cite{moore2004}.
-In this article, we will show how to design a pitch-independent timbre feature which is well-adapted to musical data, by deriving it from the constant-Q transform (CQT) \cite{brown1991}, \cite{brown1992}, a log-scaled frequency transform which matches the notes of the Western music scale.
-We will show how to decompose the magnitude CQT spectrum into an energy-normalized pitch component and a pitch-independent spectral envelope, the latter from which we will extract a number of timbral coefficients.
+In this article, we will show how to design a simple but practical pitch-independent timbre feature which is well-adapted to musical data, by deriving it from the constant-Q transform (CQT) \cite{brown1991, brown1992}, a log-scaled frequency transform which matches the notes of the Western music scale.
+We will show how to decompose the power CQT spectrum into an energy-normalized pitch component and a pitch-independent spectral envelope, the latter from which we will extract a number of timbral coefficients.
 We will then evaluate the discriminative power of these CQT spectral envelope coefficients (CQT-SEC) on the NSynth dataset \cite{engel2017}, a large-scale dataset of musical notes which is publicly available, comparing them with the mel-frequency cepstral coefficients (MFCCs) \cite{davis1980}, features originally designed for speech recognition but commonly used to characterize timbre in music.
@@
 
-A timbre feature which is well-adapted to musical data, pitch-independent, and with high discriminative power can find uses in a number of applications, such as similarity detection, sound recognition, and music classification, in particular, of musical instruments.
-Additionally, the ability to decompose the spectrum of a sound (here, the magnitude CQT) into a pitch-independent spectral envelope and an energy-normalized pitch component can be useful for audio analysis, transformation, and resynthesis.
+A timbre feature which is well-adapted to musical data, pitch-independent, and with high discriminative power can find uses in a number of applications, such as similarity detection, sound recognition, and audio classification, in particular, of musical instruments.
+Additionally, the ability to decompose the spectrum of a sound (here, the power CQT) into a pitch-independent spectral envelope and an energy-normalized pitch component can be useful for audio analysis, transformation, and resynthesis.
 The energy-normalized pitch component can also potentially be used for pitch identification and melody extraction.
@@
 
-The complex nature of timbre makes it an attribute that is not easy to quantify in terms of one single characteristic feature.
-While it is assumed to be independent from pitch and loudness, it is not really possible to fully disentangle timbre from those qualities, as timbre is inherently dependent on the spectral content of the sound, which is also defined by its pitch and loudness \cite{moore2004}.
-Researchers in MIR proposed a number of descriptors to characterize one or more aspects of timbre, but ...
-mostly adopted the MFCCs
+The multidimensional nature of timbre makes it an attribute that is tricky to quantify in terms of one single characteristic feature \cite{grey1977}.
+While it is assumed to be independent from pitch and loudness, it is not really feasible to fully disentangle timbre from those qualities, as timbre is inherently dependent on the spectral content of the sound, which is also defined by its pitch and loudness \cite{moore2004}.
+Researchers in MIR proposed a number of descriptors to characterize one or more aspects of timbre \cite{peeters2003}, but mostly adopted the MFCCs from the speech recognition community.
+While the MFCCs have shown
 
-%Given a lack of proper timbre feature for characterizing musical data, the MIR community adopted the MFCCs,
+%More recently, a number of data-driven approaches are attempting to learn some timbre representation, but can be some embeddings used for music resynthesis and that are not necessarily easily interpretable, we will show that we can use simple signal processing techniques to achieve ...
+typically, by using autoencoders and generative models
+%
+% style transfer, music generation,
 
-% MIR tried some feature linked to timbre but nothing that emcompasses the whole attribute.
-Adopted MFCCs from speech recognition, commonly-used.
-While more recent data-driven approaches are attempting to learn a timbre feature
+\cite{engel2017}
 

### 2021-05-20 IEEE SP Magazine/05-2021 - Draft/article.tex
Made little changes to the notebook and working on the solution section in the article.

@@
 In this article, we will show how to design a simple but practical pitch-independent timbre feature which is well-adapted to musical data, by deriving it from the constant-Q transform (CQT) \cite{brown1991, brown1992}, a log-scaled frequency transform which matches the notes of the Western music scale.
-We will show how to decompose the power CQT spectrum into an energy-normalized pitch component and a pitch-independent spectral envelope, the latter from which we will extract a number of timbral coefficients.
+We will show how to decompose the CQT spectrum into an energy-normalized pitch component and a pitch-independent spectral envelope, the latter from which we will extract a number of timbral coefficients.
 We will then evaluate the discriminative power of these CQT spectral envelope coefficients (CQT-SEC) on the NSynth dataset \cite{engel2017}, a large-scale dataset of musical notes which is publicly available, comparing them with the mel-frequency cepstral coefficients (MFCCs) \cite{davis1980}, features originally designed for speech recognition but commonly used to characterize timbre in music.
@@
 A timbre feature which is well-adapted to musical data, pitch-independent, and with high discriminative power can find uses in a number of applications, such as similarity detection, sound recognition, and audio classification, in particular, of musical instruments.
-Additionally, the ability to decompose the spectrum of a sound (here, the power CQT) into a pitch-independent spectral envelope and an energy-normalized pitch component can be useful for audio analysis, transformation, and resynthesis.
+Additionally, the ability to decompose the spectrum of a sound (here, the CQT spectrum) into a pitch-independent spectral envelope and an energy-normalized pitch component can be useful for audio analysis, transformation, and resynthesis.
 The energy-normalized pitch component can also potentially be used for pitch identification and melody extraction.
@@
 
-Basic knowledge of audio signal processing and some knowledge of music information retrieval (MIR) are required to understand this article, in particular, concepts such as the Fourier transform, convolution, spectral envelope, pitch, CQT, and MFCCs.
+Basic knowledge of audio signal processing and some knowledge of music information retrieval (MIR) are required to understand this article, in particular, concepts such as the Fourier transform (FT), convolution, spectral envelope, pitch, CQT, and MFCCs.
 
@@
 While it is assumed to be independent from pitch and loudness, it is not really feasible to fully disentangle timbre from those qualities, as timbre is inherently dependent on the spectral content of the sound, which is also defined by its pitch and loudness \cite{moore2004}.
-Researchers in MIR proposed a number of descriptors to characterize one or more aspects of timbre \cite{peeters2003}, but mostly adopted the MFCCs from the speech recognition community.
-While the MFCCs have shown
+While researchers in MIR proposed a number of descriptors to characterize one or more aspects of timbre \cite{peeters2003}, they mostly adopted the MFCCs as the go-to single timbre feature.
+
+% from the speech recognition community. While the MFCCs have shown
 
@@
 
-convolution theorem: \cite{proakis1995}.
+We start with the assumption that a log-spectrum $X$, in particular, the CQT spectrum, can be represented as the convolution of an energy-normalized pitch component $P$ (which mostly contains the pitch information) and a pitch-independent spectral envelope component $E$ (which mostly contains the timbre information), as shown in Equation \ref{eq:assumption}.
+\begin{equation}
+\label{eq:assumption}
+X = E * P
+\end{equation}
 
-\subsection{Observations}
+\emph{Property 1: A pitch change in the audio translates to a linear shift in the log-spectrum \cite{brown1991, brown1992}.}
 
-Assumption: A log-spectrum, such as the CQT-spectrum, can be represented as the convolution of a pitch-invariant log-specrtal envelope component (= timbre) and a envelope-normalized pitch component.
+Assuming that pitch and timbre are independent, this implies that the same musical object at different pitches would have a similar envelope component but a shifted pitch component (while two different musical objects at the same pitch would have different envelope components but a similar pitch component).
+Assuming $X'$, $E'$ and $P'$ as the log-spectrum, envelope, and pitch component of the
 
-\begin{itemize}
-\item A pitch change in the audio translates to a linear shift in the log-spectrum.
-\item The Fourier transform (FT) of a convolution of two functions is equal to the point-wise product of their FTs (convolution theorem).
-\item The magnitude FT is shift-invariant.
-\end{itemize}
+%We then point out to a first property that
+
+%  Theorem, Lemma, Proposition, Corollary, Postulate, Statement, Fact, Observation, Expression, Fact, Property, Conjecture and Principle
+
+Equation \ref{eq:property1}.
+\begin{equation}
+\label{eq:property1}
+\begin{split}
+&X = E * P \\
+&X' = E' * P' \\
+&\Rightarrow E \approx E'
+\end{split}
+\end{equation}
+
+We show in a second property that the FT of a convolution of two functions is equal to the point-wise product of their FTs, also known as the convolution theorem \cite{proakis1995}.
+
+Equation \ref{eq:property2}.
+\begin{equation}
+\label{eq:property2}
+\begin{split}
+&\mathcal{F}(X) = \mathcal{F}(E) \cdot \mathcal{F}(P) \\
+&\mathcal{F}(X') = \mathcal{F}(E') \cdot \mathcal{F}(P') \\
+&\Rightarrow \mathcal{F}(E) \approx \mathcal{F}(E')
+\end{split}
+\end{equation}
+
+in a third property that the magnitude FT is shift-invariant \cite{proakis1995}.
+
+Equation \ref{eq:property3}.
+\begin{equation}
+\label{eq:property3}
+\begin{split}
+&\mathcal{F}(X) = |\mathcal{F}(X)| \cdot e^{j Arg(\mathcal{F}(X))} \\
+&\mathcal{F}(X) = |\mathcal{F}(X')| \cdot e^{j Arg(\mathcal{F}(X'))} \\
+&\Rightarrow |\mathcal{F}(X)| \approx |\mathcal{F}(X')|
+\end{split}
+\end{equation}
+
+Given the ...,
+
+Equation \ref{eq:conclusion}.
+\begin{equation}
+\label{eq:conclusion}
+\begin{split}
+&\mathcal{F}(E) \approx |\mathcal{F}(X)| \\
+&\Rightarrow E \approx \mathcal{F}^{-1}(|\mathcal{F}(X)|) \\
+&\Rightarrow P \approx \mathcal{F}^{-1}(e^{j Arg(\mathcal{F}(X))})
+\end{split}
+\end{equation}
 
@@
 
-\section{Numerical Example}
-% or "V - Computational Example"
-
-%Equation \ref{eq:sdft} shows ...
-%
-%\begin{equation}
-%\label{eq:sdft}
-%\begin{split}
-%\underset{0 \leq k < N}{X_k^{(i)}} 
-%& = \sum_{n=0}^{N-1} x_{i+n} e^{\frac{-j 2\pi n k}{N}}\\
-%& = \sum_{n=0}^{N-1} x_{i+n} e^{\frac{-j 2\pi (n+1) k}{N}} e^{\frac{j 2\pi k}{N}} 
-%\end{split}
-%\end{equation}
+\section{Numerical Example} % or "V - Computational Example"
 
@@
 \section{What We Have Learned}
-% One paragraph
 

### 2021-05-25 IEEE SP Magazine/05-2021 - Draft/article.tex
Added more references, still working on the problem section.

@@
 It is generally associated with the spectral (but also temporal) envelope and is typically assumed to be independent from the pitch (but also the loudness) of the sound \cite{moore2004}.
-In this article, we will show how to design a simple but practical pitch-independent timbre feature which is well-adapted to musical data, by deriving it from the constant-Q transform (CQT) \cite{brown1991, brown1992}, a log-scaled frequency transform which matches the notes of the Western music scale.
+In this article, we will show how to design a simple but functional timbre feature which is well-adapted to musical data, by deriving it from the constant-Q transform (CQT) \cite{brown1991, brown1992}, a log-scaled frequency transform which matches the notes of the Western music scale.
 We will show how to decompose the CQT spectrum into an energy-normalized pitch component and a pitch-independent spectral envelope, the latter from which we will extract a number of timbral coefficients.
@@
 
-Basic knowledge of audio signal processing and some knowledge of music information retrieval (MIR) are required to understand this article, in particular, concepts such as the Fourier transform (FT), convolution, spectral envelope, pitch, CQT, and MFCCs.
+Basic knowledge of audio signal processing and some knowledge of music information retrieval (MIR) \cite{mueller2007} are required to understand this article, in particular, concepts such as the Fourier transform (FT), convolution, spectral envelope, pitch, CQT, and MFCCs.
 
@@
 The multidimensional nature of timbre makes it an attribute that is tricky to quantify in terms of one single characteristic feature \cite{grey1977}.
-While it is assumed to be independent from pitch and loudness, it is not really feasible to fully disentangle timbre from those qualities, as timbre is inherently dependent on the spectral content of the sound, which is also defined by its pitch and loudness \cite{moore2004}.
-While researchers in MIR proposed a number of descriptors to characterize one or more aspects of timbre \cite{peeters2003}, they mostly adopted the MFCCs as the go-to single timbre feature.
-
-% from the speech recognition community. While the MFCCs have shown
+While it is assumed to be independent from pitch and loudness, it is not really feasible to fully disentangle timbre from those qualities, as timbre is inherently dependent on the spectral content of the sound which is also defined by its pitch and loudness \cite{moore2004}.
+Researchers in MIR proposed a number of descriptors to characterize one or more aspects of timbre \cite{peeters2003}, but they mostly resort to using the MFCCs when it comes to having one single timbre feature \cite{mueller2007}.
+While the MFCCs were shown to be helpful in a number of MIR tasks, they were originally designed for speech processing applications \cite{davis1980}, using the mel scale, an old perceptual scale built to approximate the auditory system's response \cite{stevens1937}, and are not necessarily adapted to musical data.
 

### 2021-05-26 IEEE SP Magazine/05-2021 - Draft/article.tex
Little update.
Finished Problem section. Added new references.

@@
 We will show how to decompose the CQT spectrum into an energy-normalized pitch component and a pitch-independent spectral envelope, the latter from which we will extract a number of timbral coefficients.
-We will then evaluate the discriminative power of these CQT spectral envelope coefficients (CQT-SEC) on the NSynth dataset \cite{engel2017}, a large-scale dataset of musical notes which is publicly available, comparing them with the mel-frequency cepstral coefficients (MFCCs) \cite{davis1980}, features originally designed for speech recognition but commonly used to characterize timbre in music.
+We will then evaluate the discriminative power of these CQT spectral envelope coefficients (CQT-SEC) on the NSynth dataset \cite{engel2017}, a large-scale dataset of musical notes which is publicly available, comparing them with the mel-frequency cepstral coefficients (MFCCs) \cite{mermelstein1976}, features originally designed for speech recognition but commonly used to characterize timbre in music.
 
@@
 While it is assumed to be independent from pitch and loudness, it is not really feasible to fully disentangle timbre from those qualities, as timbre is inherently dependent on the spectral content of the sound which is also defined by its pitch and loudness \cite{moore2004}.
-Researchers in MIR proposed a number of descriptors to characterize one or more aspects of timbre \cite{peeters2003}, but they mostly resort to using the MFCCs when it comes to having one single timbre feature \cite{mueller2007}.
-While the MFCCs were shown to be helpful in a number of MIR tasks, they were originally designed for speech processing applications \cite{davis1980}, using the mel scale, an old perceptual scale built to approximate the auditory system's response \cite{stevens1937}, and are not necessarily adapted to musical data.
-
-%More recently, a number of data-driven approaches are attempting to learn some timbre representation, but can be some embeddings used for music resynthesis and that are not necessarily easily interpretable, we will show that we can use simple signal processing techniques to achieve ...
-typically, by using autoencoders and generative models
-%
-% style transfer, music generation,
-
-\cite{engel2017}
+Researchers in MIR proposed a number of descriptors to characterize one or more aspects of timbre \cite{peeters2011}, but they mostly resort to using the MFCCs when it comes to having one single timbre feature \cite{mueller2007}.
+While the MFCCs were shown to be helpful in a number of MIR tasks, they were initially designed for speech processing applications \cite{mermelstein1976} and are not necessarily adapted to musical data.
+In particular, they are computed through an old process which makes use of the mel scale, a perceptual scale experimentally developed 80 years ago to approximate the human auditory system's response \cite{stevens1937}.
+More recently, a number of data-driven approaches attempted to learn some timbral representations from musical data, but generally in terms of implicit embeddings which are tied to a specific trained model \cite{engel2017}, \cite{pons2017}, and not necessarily as in explicit and interpretable features such as the MFCCs, which are still usually preferred as the go-to single feature to characterize timbre in music.
 
 \section{Solution}
+
+%We proposed ...
+%
+%simple to compute
+%interpretable
+%functional
+%
+%We do not claim that it is perfect, but it's a better approximation
 

### 2021-05-27 IEEE SP Magazine/05-2021 - Draft/article.tex
Working on the solution section.

@@
 It is generally associated with the spectral (but also temporal) envelope and is typically assumed to be independent from the pitch (but also the loudness) of the sound \cite{moore2004}.
-In this article, we will show how to design a simple but functional timbre feature which is well-adapted to musical data, by deriving it from the constant-Q transform (CQT) \cite{brown1991, brown1992}, a log-scaled frequency transform which matches the notes of the Western music scale.
+In this article, we will show how to design a simple but functional timbre feature which is well-adapted to musical data, by deriving it from the constant-Q transform (CQT) \cite{brown1991, brown1992}, a log-scaled frequency transform which matches the Western music scale.
 We will show how to decompose the CQT spectrum into an energy-normalized pitch component and a pitch-independent spectral envelope, the latter from which we will extract a number of timbral coefficients.
@@
 Basic knowledge of audio signal processing and some knowledge of music information retrieval (MIR) \cite{mueller2007} are required to understand this article, in particular, concepts such as the Fourier transform (FT), convolution, spectral envelope, pitch, CQT, and MFCCs.
+In particular, more information about the CQT can be found in\cite{brown1991, brown1992}.
 
@@
 
-The multidimensional nature of timbre makes it an attribute that is tricky to quantify in terms of one single characteristic feature \cite{grey1977}.
-While it is assumed to be independent from pitch and loudness, it is not really feasible to fully disentangle timbre from those qualities, as timbre is inherently dependent on the spectral content of the sound which is also defined by its pitch and loudness \cite{moore2004}.
-Researchers in MIR proposed a number of descriptors to characterize one or more aspects of timbre \cite{peeters2011}, but they mostly resort to using the MFCCs when it comes to having one single timbre feature \cite{mueller2007}.
-While the MFCCs were shown to be helpful in a number of MIR tasks, they were initially designed for speech processing applications \cite{mermelstein1976} and are not necessarily adapted to musical data.
-In particular, they are computed through an old process which makes use of the mel scale, a perceptual scale experimentally developed 80 years ago to approximate the human auditory system's response \cite{stevens1937}.
-More recently, a number of data-driven approaches attempted to learn some timbral representations from musical data, but generally in terms of implicit embeddings which are tied to a specific trained model \cite{engel2017}, \cite{pons2017}, and not necessarily as in explicit and interpretable features such as the MFCCs, which are still usually preferred as the go-to single feature to characterize timbre in music.
+The multidimensional nature of timbre makes it an attribute that is tricky to quantify in terms of one simple characteristic feature \cite{grey1977}.
+While it is assumed to be independent from pitch and loudness, it is not really feasible to fully disentangle timbre from those qualities, as timbre is inherently dependent on the spectral content of the sound, which is also defined by its pitch and loudness \cite{moore2004}.
+Researchers in MIR proposed a number of descriptors to characterize one or more aspects of timbre \cite{peeters2011}, but they mostly resort to using the MFCCs when they need one simple timbre feature \cite{mueller2007}.
+While the MFCCs were shown to be helpful in some MIR tasks, they were initially designed for speech processing applications \cite{mermelstein1976} and are not necessarily adapted to musical data.
+In particular, they are derived through an old process which makes use of the mel scale, a perceptual scale experimentally designed 80 years ago to approximate the human auditory system's response \cite{stevens1937}.
+More recently, a number of data-driven approaches attempted to learn some timbral representations from musical data, but generally in terms of implicit embeddings which are tied to a specific trained model \cite{engel2017}, \cite{pons2017}, and not necessarily as explicit and interpretable features such as the MFCCs, which are still usually preferred as the go-to feature to characterize musical timbre by MIR practitioners.
 
@@
 
-%We proposed ...
-%
-%simple to compute
-%interpretable
-%functional
-%
-%We do not claim that it is perfect, but it's a better approximation
+We propose the CQT-SECs, a novel timbre feature that is well-adapted to musical data, pitch-independent, simple to derive, interpretable, and functional.
+We will show how to derive it from the CQT, a log-frequency transform which matches the Western music scale, by first decomposing it into a pitch-independent spectral envelope and an energy-normalized pitch component, and then extracting a number of timbral coefficients from the spectral envelope.
 
-%- The envelope component can additionally be refined, along with the pitch component
-%- Incidentally, the pitch component can be used to identify the pitch/key
-
-We start with the assumption that a log-spectrum $X$, in particular, the CQT spectrum, can be represented as the convolution of an energy-normalized pitch component $P$ (which mostly contains the pitch information) and a pitch-independent spectral envelope component $E$ (which mostly contains the timbre information), as shown in Equation \ref{eq:assumption}.
+We start with the assumption that a log-spectrum $X$, in particular, the CQT spectrum, can be represented as the convolution between a pitch-independent spectral envelope component $E$ (which mostly contains the timbre information) and an energy-normalized pitch component $P$ (which mostly contains the pitch information), as shown in Equation \ref{eq:assumption}.
 \begin{equation}
@@
 
-\emph{Property 1: A pitch change in the audio translates to a linear shift in the log-spectrum \cite{brown1991, brown1992}.}
-
-Assuming that pitch and timbre are independent, this implies that the same musical object at different pitches would have a similar envelope component but a shifted pitch component (while two different musical objects at the same pitch would have different envelope components but a similar pitch component).
-Assuming $X'$, $E'$ and $P'$ as the log-spectrum, envelope, and pitch component of the
-
-%We then point out to a first property that
-
-%  Theorem, Lemma, Proposition, Corollary, Postulate, Statement, Fact, Observation, Expression, Fact, Property, Conjecture and Principle
-
-Equation \ref{eq:property1}.
+We make a first observation that a pitch change in the audio translates to a linear shift in the log-spectrum \cite{brown1991, brown1992}.
+Assuming that pitch and timbre are independent, this implies that the same musical object at different pitches would have a similar envelope but a shifted pitch component (while two different musical objects at the same pitch would have different envelopes but a similar pitch component).
+Assuming $X$, $E$, $P$ and $X'$, $E'$, $P'$ as the log-spectrum, envelope, and pitch component of a musical object and a pitch-shifted version of it, respectively, this is summarized in equation \ref{eq:observation1}.
 \begin{equation}
-\label{eq:property1}
+\label{eq:observation1}
 \begin{split}
@@
 
-We show in a second property that the FT of a convolution of two functions is equal to the point-wise product of their FTs, also known as the convolution theorem \cite{proakis1995}.
-
-Equation \ref{eq:property2}.
+We make a second observation that the FT of a convolution of two functions is equal to the point-wise product of their FTs, also known as the convolution theorem \cite{proakis1995}.
+This implies that the FT of the log-spectrum is equal to the point-wise product of the FT of the envelope and the FT of the pitch component.
+Given the previous observation that the envelope is not affected by a pitch change in the audio, we conclude that the FT of the envelope is also independent from the pitch.
+Assuming that $\mathcal{F}(.)$ is the FT function, this is summarized in equation \ref{eq:observation2}.
 \begin{equation}
-\label{eq:property2}
+\label{eq:observation2}
 \begin{split}
@@
 
-in a third property that the magnitude FT is shift-invariant \cite{proakis1995}.
+We make a final and third observation that the magnitude FT is shift-invariant \cite{proakis1995}. ...
 
-Equation \ref{eq:property3}.
+Equation \ref{eq:observation3}.
 \begin{equation}
-\label{eq:property3}
+\label{eq:observation3}
 \begin{split}
@@
 % Can think about the extraction of the envelop as a white normalization of the pitch by using a weighting function that normalizes the signal spectral density by the spectrum magnitude, just like in the GCC-PHAT
+
+%- The envelope component can additionally be refined, along with the pitch component
+%- Incidentally, the pitch component can be used to identify the pitch/key
 

### 2021-05-28 IEEE SP Magazine/05-2021 - Draft/article.tex
Updated the solution section.

@@
 Assuming that pitch and timbre are independent, this implies that the same musical object at different pitches would have a similar envelope but a shifted pitch component (while two different musical objects at the same pitch would have different envelopes but a similar pitch component).
-Assuming $X$, $E$, $P$ and $X'$, $E'$, $P'$ as the log-spectrum, envelope, and pitch component of a musical object and a pitch-shifted version of it, respectively, this is summarized in equation \ref{eq:observation1}.
+This is summarized in equation \ref{eq:observation1}.
 \begin{equation}
@@
 \end{equation}
+where $X$, $E$, $P$ and $X'$, $E'$, $P'$ represent the log-spectrum, envelope, and pitch component of a musical object and a pitch-shifted version of it, respectively.
 
-We make a second observation that the FT of a convolution of two functions is equal to the point-wise product of their FTs, also known as the convolution theorem \cite{proakis1995}.
-This implies that the FT of the log-spectrum is equal to the point-wise product of the FT of the envelope and the FT of the pitch component.
-Given the previous observation that the envelope is not affected by a pitch change in the audio, we conclude that the FT of the envelope is also independent from the pitch.
-Assuming that $\mathcal{F}(.)$ is the FT function, this is summarized in equation \ref{eq:observation2}.
+We make a second observation that the FT of a convolution of two functions is equal to the point-wise product of their FTs, which is also known as the convolution theorem \cite{proakis1995}.
+This implies that the FT of the log-spectrum would be equal to the point-wise product of the FT of the envelope and the FT of the pitch component.
+Given the previous observation, we infer that the FT of the envelope for a musical object and a pitch-shifted version of it would be equal.
+This is summarized in equation \ref{eq:observation2}.
 \begin{equation}
@@
 \end{equation}
+where $\mathcal{F}(.)$ represents the FT function.
 
-We make a final and third observation that the magnitude FT is shift-invariant \cite{proakis1995}. ...
+We make a final and third observation that the magnitude FT is shift-invariant \cite{proakis1995}. This implies that
 
@@
 &\mathcal{F}(X) = |\mathcal{F}(X)| \cdot e^{j Arg(\mathcal{F}(X))} \\
-&\mathcal{F}(X) = |\mathcal{F}(X')| \cdot e^{j Arg(\mathcal{F}(X'))} \\
+&\mathcal{F}(X') = |\mathcal{F}(X')| \cdot e^{j Arg(\mathcal{F}(X'))} \\
 &\Rightarrow |\mathcal{F}(X)| \approx |\mathcal{F}(X')|
@@
 \end{equation}
+where $|.|$ and $Arg(.)$ represent the modulus and argument, respectively, for a complex array.
 

### 2021-05-28 IEEE SP Magazine/05-2021 - Draft/article.tex
Still working on the solution section.

@@
 It is generally associated with the spectral (but also temporal) envelope and is typically assumed to be independent from the pitch (but also the loudness) of the sound \cite{moore2004}.
-In this article, we will show how to design a simple but functional timbre feature which is well-adapted to musical data, by deriving it from the constant-Q transform (CQT) \cite{brown1991, brown1992}, a log-scaled frequency transform which matches the Western music scale.
+In this article, we will show how to design a simple but functional timbre feature which is well-adapted to musical data, by deriving it from the constant-Q transform (CQT) \cite{brown1991, brown1992}, a log-scale frequency transform which matches the Western music scale.
 We will show how to decompose the CQT spectrum into an energy-normalized pitch component and a pitch-independent spectral envelope, the latter from which we will extract a number of timbral coefficients.
@@
 A timbre feature which is well-adapted to musical data, pitch-independent, and with high discriminative power can find uses in a number of applications, such as similarity detection, sound recognition, and audio classification, in particular, of musical instruments.
-Additionally, the ability to decompose the spectrum of a sound (here, the CQT spectrum) into a pitch-independent spectral envelope and an energy-normalized pitch component can be useful for audio analysis, transformation, and resynthesis.
+Additionally, the ability to decompose the spectrum of a sound (here, the CQT spectrum) into a pitch-independent spectral envelope and an energy-normalized pitch component can be useful for analysis, transformation, and resynthesis of music signals.
 The energy-normalized pitch component can also potentially be used for pitch identification and melody extraction.
@@
 Basic knowledge of audio signal processing and some knowledge of music information retrieval (MIR) \cite{mueller2007} are required to understand this article, in particular, concepts such as the Fourier transform (FT), convolution, spectral envelope, pitch, CQT, and MFCCs.
-In particular, more information about the CQT can be found in\cite{brown1991, brown1992}.
+More information about the CQT can also be found in\cite{brown1991, brown1992}.
 
@@
 
-We start with the assumption that a log-spectrum $X$, in particular, the CQT spectrum, can be represented as the convolution between a pitch-independent spectral envelope component $E$ (which mostly contains the timbre information) and an energy-normalized pitch component $P$ (which mostly contains the pitch information), as shown in Equation \ref{eq:assumption}.
+\subsection{Assumption and Observations}
+
+We start with the assumption that a log-spectrum $X$, in particular, the CQT spectrum, can be represented as the convolution between a pitch-independent spectral envelope $E$ (which mostly contains the timbre information) and an energy-normalized pitch component $P$ (which mostly contains the pitch information), as shown in Equation \ref{eq:assumption}.
 \begin{equation}
@@
 
-We make a first observation that a pitch change in the audio translates to a linear shift in the log-spectrum \cite{brown1991, brown1992}.
+We then make a first observation that a pitch change in the audio translates to a linear shift in the log-spectrum \cite{brown1991, brown1992}.
 Assuming that pitch and timbre are independent, this implies that the same musical object at different pitches would have a similar envelope but a shifted pitch component (while two different musical objects at the same pitch would have different envelopes but a similar pitch component).
-This is summarized in equation \ref{eq:observation1}.
+This is summarized in Equation \ref{eq:observation1}.
 \begin{equation}
@@
 \begin{split}
-&X = E * P \\
-&X' = E' * P' \\
-&\Rightarrow E \approx E'
+\begin{cases}
+X = E * P \\
+X' = E' * P' \\
+\end{cases}
+\Rightarrow E \approx E'
 \end{split}
 \end{equation}
-where $X$, $E$, $P$ and $X'$, $E'$, $P'$ represent the log-spectrum, envelope, and pitch component of a musical object and a pitch-shifted version of it, respectively.
+where $X$, $E$, $P$ and $X'$, $E'$, $P'$ represent the log-spectrum, envelope, and pitch component for a musical object and for a pitch-shifted version of it, respectively.
 
@@
 This implies that the FT of the log-spectrum would be equal to the point-wise product of the FT of the envelope and the FT of the pitch component.
-Given the previous observation, we infer that the FT of the envelope for a musical object and a pitch-shifted version of it would be equal.
+Given the previous observation, this also implies that the FT of the envelope for a musical object and for a pitch-shifted version of it would be equal.
 This is summarized in equation \ref{eq:observation2}.
@@
 \begin{split}
-&\mathcal{F}(X) = \mathcal{F}(E) \cdot \mathcal{F}(P) \\
-&\mathcal{F}(X') = \mathcal{F}(E') \cdot \mathcal{F}(P') \\
-&\Rightarrow \mathcal{F}(E) \approx \mathcal{F}(E')
+\begin{cases}
+\mathcal{F}(X) = \mathcal{F}(E) \cdot \mathcal{F}(P) \\
+\mathcal{F}(X') = \mathcal{F}(E') \cdot \mathcal{F}(P') \\
+\end{cases}
+\Rightarrow \mathcal{F}(E) \approx \mathcal{F}(E')
 \end{split}
@@
 
-We make a final and third observation that the magnitude FT is shift-invariant \cite{proakis1995}. This implies that
+We make a third and final observation that the magnitude FT is shift-invariant \cite{proakis1995}. This implies that the FT of the log-spectrum is equal
 
@@
 \begin{split}
-&\mathcal{F}(X) = |\mathcal{F}(X)| \cdot e^{j Arg(\mathcal{F}(X))} \\
-&\mathcal{F}(X') = |\mathcal{F}(X')| \cdot e^{j Arg(\mathcal{F}(X'))} \\
-&\Rightarrow |\mathcal{F}(X)| \approx |\mathcal{F}(X')|
+\begin{cases}
+\mathcal{F}(X) = |\mathcal{F}(X)| \cdot e^{j Arg(\mathcal{F}(X))} \\
+\mathcal{F}(X') = |\mathcal{F}(X')| \cdot e^{j Arg(\mathcal{F}(X'))} \\
+\end{cases}
+\Rightarrow |\mathcal{F}(X)| \approx |\mathcal{F}(X')|
 \end{split}
@@
 \end{equation}
+
+\subsection{Derivation}
 

### 2021-05-29 IEEE SP Magazine/05-2021 - Draft/article.tex
Finished first part of solution section.

@@
 Researchers in MIR proposed a number of descriptors to characterize one or more aspects of timbre \cite{peeters2011}, but they mostly resort to using the MFCCs when they need one simple timbre feature \cite{mueller2007}.
-While the MFCCs were shown to be helpful in some MIR tasks, they were initially designed for speech processing applications \cite{mermelstein1976} and are not necessarily adapted to musical data.
-In particular, they are derived through an old process which makes use of the mel scale, a perceptual scale experimentally designed 80 years ago to approximate the human auditory system's response \cite{stevens1937}.
+While the MFCCs were shown to be helpful in some MIR tasks, they were initially designed for speech processing applications \cite{mermelstein1976} and are not necessarily well-adapted to musical data.
+In particular, they are derived through an old process which makes use of the mel scale, a perceptual scale experimentally designed 85 years ago to approximate the human auditory system's response \cite{stevens1937}.
 More recently, a number of data-driven approaches attempted to learn some timbral representations from musical data, but generally in terms of implicit embeddings which are tied to a specific trained model \cite{engel2017}, \cite{pons2017}, and not necessarily as explicit and interpretable features such as the MFCCs, which are still usually preferred as the go-to feature to characterize musical timbre by MIR practitioners.
@@
 
-We propose the CQT-SECs, a novel timbre feature that is well-adapted to musical data, pitch-independent, simple to derive, interpretable, and functional.
+%We propose the CQT-SECs, a novel timbre feature that is well-adapted to musical data, pitch-independent, simple to derive, interpretable, and functional.
 We will show how to derive it from the CQT, a log-frequency transform which matches the Western music scale, by first decomposing it into a pitch-independent spectral envelope and an energy-normalized pitch component, and then extracting a number of timbral coefficients from the spectral envelope.
 
-\subsection{Assumption and Observations}
+% We will make an assumption, list 3 observations, and give a conclusion...
+
+\subsection{Deconvolution of the CQT}
 
@@
 
-We then make a first observation that a pitch change in the audio translates to a linear shift in the log-spectrum \cite{brown1991, brown1992}.
+This convolution process can be thought of a source-filter model \cite{fant1970} which is here not applied in the time domain but the frequency domain, with the source and the filter being the pitch component and the envelope, respectively.
+
+\emph{Observation 1: A pitch change in the audio translates to a linear shift in the log-spectrum \cite{brown1991, brown1992}.}
+
 Assuming that pitch and timbre are independent, this implies that the same musical object at different pitches would have a similar envelope but a shifted pitch component (while two different musical objects at the same pitch would have different envelopes but a similar pitch component).
-This is summarized in Equation \ref{eq:observation1}.
+This is summarized in Equation \ref{eq:observation1}, where $X$, $E$, $P$ and $X'$, $E'$, $P'$ represent the log-spectrum, envelope, and pitch component for a musical object and for a pitch-shifted version of it, respectively.
 \begin{equation}
@@
 \end{equation}
-where $X$, $E$, $P$ and $X'$, $E'$, $P'$ represent the log-spectrum, envelope, and pitch component for a musical object and for a pitch-shifted version of it, respectively.
 
-We make a second observation that the FT of a convolution of two functions is equal to the point-wise product of their FTs, which is also known as the convolution theorem \cite{proakis1995}.
-This implies that the FT of the log-spectrum would be equal to the point-wise product of the FT of the envelope and the FT of the pitch component.
+\emph{Observation 2: the FT of a convolution of two functions is equal to the point-wise product of their FTs, which is also known as the convolution theorem \cite{proakis1995}.}
+
+This implies that the FT of the log-spectrum is equal to the point-wise product of the FT of the envelope and the FT of the pitch component.
 Given the previous observation, this also implies that the FT of the envelope for a musical object and for a pitch-shifted version of it would be equal.
-This is summarized in equation \ref{eq:observation2}.
+This is summarized in Equation \ref{eq:observation2}, where $\mathcal{F}(.)$ represents the FT function.
 \begin{equation}
@@
 \end{equation}
-where $\mathcal{F}(.)$ represents the FT function.
 
-We make a third and final observation that the magnitude FT is shift-invariant \cite{proakis1995}. This implies that the FT of the log-spectrum is equal
+\emph{Observation 3: The magnitude FT is shift-invariant \cite{proakis1995}.}
 
-Equation \ref{eq:observation3}.
+This implies that the magnitude of the FT of the log-spectrum for a musical object and for a pitch-shifted version of it would be equal.
+This is summarized in Equation \ref{eq:observation3}, where $|.|$ and $Arg(.)$ represent the modulus and argument, respectively, for a complex array.
 \begin{equation}
@@
 \end{equation}
-where $|.|$ and $Arg(.)$ represent the modulus and argument, respectively, for a complex array.
 
-Given the ...,
-
-Equation \ref{eq:conclusion}.
+Given the previous observations, we can therefore conclude that the FT of the envelope could be approximated by the magnitude of FT of the log-spectrum, while the FT of the pitch component by the phase component, which finally gives us estimates for the envelope and pitch component after taking their inverse FTs, as shown in Equation \ref{eq:conclusion}, where $\mathcal{F}^{-1}(.)$ represents the inverse FT function.
 \begin{equation}
@@
 \begin{split}
-&\mathcal{F}(E) \approx |\mathcal{F}(X)| \\
-&\Rightarrow E \approx \mathcal{F}^{-1}(|\mathcal{F}(X)|) \\
-&\Rightarrow P \approx \mathcal{F}^{-1}(e^{j Arg(\mathcal{F}(X))})
+\Rightarrow 
+\begin{cases}
+\mathcal{F}(E) \approx |\mathcal{F}(X)| &\Rightarrow E \approx \mathcal{F}^{-1}(|\mathcal{F}(X)|) \\
+\mathcal{F}(P) \approx e^{j Arg(\mathcal{F}(X))} &\Rightarrow P \approx \mathcal{F}^{-1}(e^{j Arg(\mathcal{F}(X))}) \\
+\end{cases}
 \end{split}
@@
 
-\subsection{Derivation}
+This deconvolution process can also be thought of the normalization of the log-spectrum by the magnitude of its FT (which here corresponds to the FT of the envelope) leading to a sharper log-spectrum (which here corresponds to the pitch component), in the manner of the generalized cross-correlation phase transform (GCC-PHAT) method which aims at normalizing a cross-correlation function by its magnitude spectrum to sharpen the cross-correlation peaks \cite{knapp1976}.
 
-% Can think about the extraction of the envelop as a white normalization of the pitch by using a weighting function that normalizes the signal spectral density by the spectrum magnitude, just like in the GCC-PHAT
+\subsection{Extraction of the Coefficients}
+
+Once the CQT spectrum has been decomposed into its envelope and pitch component, ...
+
+\section{Numerical Examples} % or "V - Computational Example"
+
+\subsection{Analysis on an Example}
 
 %- The envelope component can additionally be refined, along with the pitch component
-%- Incidentally, the pitch component can be used to identify the pitch/key
 
-\section{Numerical Example} % or "V - Computational Example"
+\subsection{Comparison on a Dataset}
 
@@
 
-We have shown that ...
+We have shown that we can derive a simple but function timbre feature which is more adapted to musical data ...
 

### 2021-05-29 IEEE SP Magazine/05-2021 - Draft/article.tex
Still working on the solution section.

@@
 
-This convolution process can be thought of a source-filter model \cite{fant1970} which is here not applied in the time domain but the frequency domain, with the source and the filter being the pitch component and the envelope, respectively.
+This convolution process can be thought of a source-filter model \cite{fant1970} which is here not applied in the time domain but in the frequency domain, with the source and the filter being the pitch component and the envelope, respectively.
 
@@
 This implies that the FT of the log-spectrum is equal to the point-wise product of the FT of the envelope and the FT of the pitch component.
-Given the previous observation, this also implies that the FT of the envelope for a musical object and for a pitch-shifted version of it would be equal.
+Given the previous observation, this further implies that the FT of the envelope for a musical object and for a pitch-shifted version of it would be equal.
 This is summarized in Equation \ref{eq:observation2}, where $\mathcal{F}(.)$ represents the FT function.
@@
 
-Given the previous observations, we can therefore conclude that the FT of the envelope could be approximated by the magnitude of FT of the log-spectrum, while the FT of the pitch component by the phase component, which finally gives us estimates for the envelope and pitch component after taking their inverse FTs, as shown in Equation \ref{eq:conclusion}, where $\mathcal{F}^{-1}(.)$ represents the inverse FT function.
+Given the previous observations, we can therefore conclude that the FT of the envelope could be approximated by the magnitude of the FT of the log-spectrum, while the FT of the pitch component, by the phase component, which finally gives us estimates for the envelope and the pitch component after taking their inverse FTs, as shown in Equation \ref{eq:conclusion}, where $\mathcal{F}^{-1}(.)$ represents the inverse FT function.
 \begin{equation}
@@
 
-This deconvolution process can also be thought of the normalization of the log-spectrum by the magnitude of its FT (which here corresponds to the FT of the envelope) leading to a sharper log-spectrum (which here corresponds to the pitch component), in the manner of the generalized cross-correlation phase transform (GCC-PHAT) method which aims at normalizing a cross-correlation function by its magnitude spectrum to sharpen the cross-correlation peaks \cite{knapp1976}.
+This deconvolution process can also be thought of the normalization of the log-spectrum by the magnitude of its FT (which here would correspond to the FT of the envelope) leading to a sharper log-spectrum (which here would correspond to the pitch component), in the manner of the generalized cross-correlation phase transform (GCC-PHAT) method which aims at normalizing a cross-correlation function by its magnitude spectrum to sharpen the cross-correlation peaks \cite{knapp1976}.
 

### 2021-06-02 cqtsec.py
Started working on the Python module. For tests and to get published too.

@@ -1,0 +1,115 @@
+"""
+This Python module implements the constant-Q transform spectral envelope coefficients (CQT-SEC) and other related functions.
+
+Functions:
+    mfcc - Compute the mel-frequency cepstral coefficients (MFCCs) using librosa.
+    cqt - Compute the magnitude constant-Q transform (CQT) spectrogram using librosa.
+    cqtdeconv - Compute the pitch-independent envelope and the energy-normalized pitch component.
+
+Author:
+    Zafar Rafii
+    zafarrafii@gmail.com
+    http://zafarrafii.com
+    https://github.com/zafarrafii
+    https://www.linkedin.com/in/zafarrafii/
+    06/02/21
+"""
+
+import numpy as np
+import librosa
+
+
+def mfcc(audio_signal, sampling_frequency, number_coefficients=20):
+    """
+    Compute the mel-frequency cepstral coefficients (MFCCs) using librosa.
+
+    Inputs:
+        audio_signal: audio signal (number_samples,)
+        sampling_frequency: sampling frequency in Hz
+        number_coefficients: number of MFCCs (default: 20 coefficients)
+    Output:
+        audio_mfcc: audio MFCCs (number_coefficients, number_frames)
+    """
+
+    # Set the window and step length in samples
+    window_length = pow(2, int(np.ceil(np.log2(0.04 * sampling_frequency))))
+    step_length = int(window_length / 2)
+
+    # Compute the MFCCs using librosa's mfcc
+    audio_mfcc = librosa.feature.mfcc(
+        y=audio_signal,
+        sr=sampling_frequency,
+        n_fft=window_length,
+        hop_length=step_length,
+        n_mfcc=number_coefficients,
+    )
+
+    return audio_mfcc
+
+
+def cqt(
+    audio_signal, sampling_frequency, minimum_frequency=32.70, octave_resolution=12
+):
+    """
+    Compute the magnitude constant-Q transform (CQT) spectrogram using librosa.
+
+    Inputs:
+        audio_signal: audio signal (number_samples,)
+        sampling_frequency: sampling frequency in Hz
+        minimum_frequency: minimum frequency in Hz (default: 32.70 Hz = C1)
+        octave_resolution: number of frequency channels per octave (default: 12 frequency channels per octave)
+    Output:
+        cqt_spectrogram: magnitude CQT spectrogram (number_frequencies, number_frames)
+    """
+
[…]

### 2021-06-16 cqtsec.py
Re-re-doing similarities for only velocity of 075.

@@ -5,7 +5,7 @@ Functions:
     mfcc - Compute the mel-frequency cepstral coefficients (MFCCs) using librosa.
     cqt - Compute the magnitude constant-Q transform (CQT) spectrogram using librosa.
     cqtdeconv - Compute the pitch-independent spectral envelope and the energy-normalized pitch component from the CQT spectrogram.
-    cqtsec - Extract the CQT spectral envelope coefficients (CQT-SEC) from the CQT spectral envelope.
+    cqtsec - Compute the constant-Q transform (CQT) spectral envelope coefficients (CQT-SEC).
 
 Author:
     Zafar Rafii
@@ -13,7 +13,7 @@ Author:
     http://zafarrafii.com
     https://github.com/zafarrafii
     https://www.linkedin.com/in/zafarrafii/
-    06/14/21
+    06/16/21
 """
 
 import numpy as np
@@ -120,18 +120,49 @@ def cqtdeconv(cqt_spectrogram):
     return cqt_envelope, cqt_pitch
 
 
-def cqtsec(cqt_envelope, octave_resolution=12, number_coefficients=20):
+def cqtsec(
+    audio_signal,
+    sampling_frequency,
+    step_length,
+    minimum_frequency=32.70,
+    octave_resolution=12,
+    number_coefficients=20,
+):
     """
-    Extract the constant-Q transform (CQT) spectral envelope coefficients (CQT-SEC) from the CQT spectral envelope.
+    Compute the constant-Q transform (CQT) spectral envelope coefficients (CQT-SEC).
 
     Inputs:
-        cqt_envelope: CQT spectral envelope (number_frequencies, number_frames)
+        audio_signal: audio signal (number_samples,)
+        sampling_frequency: sampling frequency in Hz
+        step_length: step length in samples
+        minimum_frequency: minimum frequency in Hz (default: 32.70 Hz = C1)
         octave_resolution: number of frequency channels per octave (default: 12 frequency channels per octave)
         number_coefficients: number of CQT-SECs (default: 20 coefficients)
     Output:
         cqt_sec: CQT-SECs (number_coefficients, number_frames)
     """
-    # Derive the indices of the CQT-SECs in the CQT spectral envelope and extract them
+
+    # Compute the power CQT spectrogram
+    cqt_spectrogram = np.power(
+        cqt(
+            audio_signal,
+            sampling_frequency,
+            step_length,
+            minimum_frequency,
[…]

### 2021-06-18 IEEE SP Magazine/06-2021 - Draft/article.tex
Updated article and notebook.

--- 2021-05-29 IEEE SP Magazine/05-2021 - Draft/article.tex
@@
 It is generally associated with the spectral (but also temporal) envelope and is typically assumed to be independent from the pitch (but also the loudness) of the sound \cite{moore2004}.
-In this article, we will show how to design a simple but functional timbre feature which is well-adapted to musical data, by deriving it from the constant-Q transform (CQT) \cite{brown1991, brown1992}, a log-scale frequency transform which matches the Western music scale.
+In this article, we will show how to design a simple but functional timbre feature which is well-adapted to musical data, by deriving it from the constant-Q transform (CQT) \cite{brown1991, brown1992}, a log-frequency transform which matches the Western music scale.
 We will show how to decompose the CQT spectrum into an energy-normalized pitch component and a pitch-independent spectral envelope, the latter from which we will extract a number of timbral coefficients.
@@
 Additionally, the ability to decompose the spectrum of a sound (here, the CQT spectrum) into a pitch-independent spectral envelope and an energy-normalized pitch component can be useful for analysis, transformation, and resynthesis of music signals.
-The energy-normalized pitch component can also potentially be used for pitch identification and melody extraction.
+The energy-normalized pitch component can also potentially be used for tasks such as pitch identification, melody extraction, and chord recognition.
 
@@
 Basic knowledge of audio signal processing and some knowledge of music information retrieval (MIR) \cite{mueller2007} are required to understand this article, in particular, concepts such as the Fourier transform (FT), convolution, spectral envelope, pitch, CQT, and MFCCs.
-More information about the CQT can also be found in\cite{brown1991, brown1992}.
+More information about the CQT can also be found in \cite{brown1991, brown1992}.
 
@@
 
-%We propose the CQT-SECs, a novel timbre feature that is well-adapted to musical data, pitch-independent, simple to derive, interpretable, and functional.
-We will show how to derive it from the CQT, a log-frequency transform which matches the Western music scale, by first decomposing it into a pitch-independent spectral envelope and an energy-normalized pitch component, and then extracting a number of timbral coefficients from the spectral envelope.
-
-% We will make an assumption, list 3 observations, and give a conclusion...
+We propose here the CQT-SECs, a novel timbre feature which is well-adapted to musical data, pitch-independent, simple to compute, interpretable, and functional.
+We will show how to derive it from the CQT, a frequency transform with a logarithmic resolution which matches the notes of Western music scale \cite{brown1991, brown1992}, by first decomposing the CQT log-spectrum into a pitch-independent spectral envelope and an energy-normalized pitch component, and then extracting a number of timbral coefficients from the spectral envelope.
 
@@
 Assuming that pitch and timbre are independent, this implies that the same musical object at different pitches would have a similar envelope but a shifted pitch component (while two different musical objects at the same pitch would have different envelopes but a similar pitch component).
-This is summarized in Equation \ref{eq:observation1}, where $X$, $E$, $P$ and $X'$, $E'$, $P'$ represent the log-spectrum, envelope, and pitch component for a musical object and for a pitch-shifted version of it, respectively.
+This is summarized in Equation \ref{eq:observation1}, where $X$, $E$, $P$ and $X'$, $E'$, $P'$ represent the log-spectrum, envelope, and pitch component for a musical object and for a pitch-shifted version of the same musical object, respectively.
 \begin{equation}
@@
 
-\emph{Observation 2: the FT of a convolution of two functions is equal to the point-wise product of their FTs, which is also known as the convolution theorem \cite{proakis1995}.}
+\emph{Observation 2: the FT of a convolution of two functions is equal to the point-wise product of their FTs, a property also known as the convolution theorem \cite{proakis1995}.}
 
 This implies that the FT of the log-spectrum is equal to the point-wise product of the FT of the envelope and the FT of the pitch component.
-Given the previous observation, this further implies that the FT of the envelope for a musical object and for a pitch-shifted version of it would be equal.
+Given the first observation, this further implies that the FT of the envelope for a musical object and for a pitch-shifted version of it would be equal.
 This is summarized in Equation \ref{eq:observation2}, where $\mathcal{F}(.)$ represents the FT function.
@@
 
-Given the previous observations, we can therefore conclude that the FT of the envelope could be approximated by the magnitude of the FT of the log-spectrum, while the FT of the pitch component, by the phase component, which finally gives us estimates for the envelope and the pitch component after taking their inverse FTs, as shown in Equation \ref{eq:conclusion}, where $\mathcal{F}^{-1}(.)$ represents the inverse FT function.
+Given the previous observations, we can therefore conclude that the FT of the envelope could be approximated by the magnitude of the FT of the log-spectrum, while the FT of the pitch component could be approximated by the phase component.
+This finally gives us the estimates for the envelope and the pitch component, after taking their inverse FTs, as shown in Equation \ref{eq:conclusion}, where $\mathcal{F}^{-1}(.)$ represents the inverse FT function.
 \begin{equation}
@@
 
-\subsection{Extraction of the Coefficients}
+\subsection{Extraction of the spectral envelope coefficients}
 
-Once the CQT spectrum has been decomposed into its envelope and pitch component, ...
+Once the CQT spectrum has been decomposed into a spectral envelope and a pitch component, we can then assumed that most of the pitch information has been removed from the envelope which now mostly contains the timbre information.
+We can think of the spectral envelope as a pitch-normalized spectrum ...
+
+Briefly explain what how the MFCCs are computed \cite{mermelstein1976}
+mel scale: \cite{stevens1937}
 

### 2021-06-22 IEEE SP Magazine/06-2021 - Draft/article.tex
Added figure and explanations.

@@
 It is generally associated with the spectral (but also temporal) envelope and is typically assumed to be independent from the pitch (but also the loudness) of the sound \cite{moore2004}.
-In this article, we will show how to design a simple but functional timbre feature which is well-adapted to musical data, by deriving it from the constant-Q transform (CQT) \cite{brown1991, brown1992}, a log-frequency transform which matches the Western music scale.
-We will show how to decompose the CQT spectrum into an energy-normalized pitch component and a pitch-independent spectral envelope, the latter from which we will extract a number of timbral coefficients.
-We will then evaluate the discriminative power of these CQT spectral envelope coefficients (CQT-SEC) on the NSynth dataset \cite{engel2017}, a large-scale dataset of musical notes which is publicly available, comparing them with the mel-frequency cepstral coefficients (MFCCs) \cite{mermelstein1976}, features originally designed for speech recognition but commonly used to characterize timbre in music.
+In this article, we will show how to design a simple but functional pitch-independent timbre feature which is well-adapted to musical data, by deriving it from the constant-Q transform (CQT) \cite{brown1991, brown1992}, a log-frequency transform which matches the Western music scale.
+We will show how to decompose the CQT-spectrum into an energy-normalized pitch component and a pitch-independent spectral envelope, the latter from which we will extract a number of timbral coefficients.
+We will then evaluate the discriminative power of these CQT-spectral envelope coefficients (CQT-SEC) on the NSynth dataset \cite{engel2017}, a large-scale dataset of musical notes which is publicly available, comparing them with the mel-frequency cepstral coefficients (MFCCs) \cite{mermelstein1976}, features originally designed for speech recognition but commonly used to characterize timbre in music.
 
@@
 A timbre feature which is well-adapted to musical data, pitch-independent, and with high discriminative power can find uses in a number of applications, such as similarity detection, sound recognition, and audio classification, in particular, of musical instruments.
-Additionally, the ability to decompose the spectrum of a sound (here, the CQT spectrum) into a pitch-independent spectral envelope and an energy-normalized pitch component can be useful for analysis, transformation, and resynthesis of music signals.
+Additionally, the ability to decompose the spectrum of a sound (here, the CQT-spectrum) into a pitch-independent spectral envelope and an energy-normalized pitch component can be useful for analysis, transformation, and resynthesis of music signals.
 The energy-normalized pitch component can also potentially be used for tasks such as pitch identification, melody extraction, and chord recognition.
@@
 
-We start with the assumption that a log-spectrum $X$, in particular, the CQT spectrum, can be represented as the convolution between a pitch-independent spectral envelope $E$ (which mostly contains the timbre information) and an energy-normalized pitch component $P$ (which mostly contains the pitch information), as shown in Equation \ref{eq:assumption}.
+We start with the assumption that a log-spectrum $X$, in particular, the CQT-spectrum, can be represented as the convolution between a pitch-independent spectral envelope $E$ (which mostly contains the timbre information) and an energy-normalized pitch component $P$ (which mostly contains the pitch information), as shown in Equation \ref{eq:assumption}.
 \begin{equation}
@@
 
-\emph{Observation 2: the FT of a convolution of two functions is equal to the point-wise product of their FTs, a property also known as the convolution theorem \cite{proakis1995}.}
+\emph{Observation 2: the FT of the convolution of two functions is equal to the point-wise product of the FTs of the two functions, a property also known as the convolution theorem \cite{proakis1995}.}
 
@@
 
+Figure \ref{fig:deconvolution} shows and example of deconvolution of the (power) CQT-spectrogram into its envelope and pitch component.
+The CQT-spectogram was computed from a 48-second audio signal created by concatenating 12 4-second acoustic bass notes, playing successively from C1 (32.70 Hz) to B1 (61.74 Hz).
+The notes come from the NSynth dataset \cite{engel2017}, for instrument \texttt{bass\_acoustic\_000}, MIDI numbers \texttt{024} to \texttt{035}, and velocity number \texttt{075}.
+As we can see, the envelope gets pitch-normalized, shifting all the frequency components to the origin of the log-spectrum (essentially bringing all the notes to C1), while the pitch component gets energy-normalized, leaving mostly the fundamental frequency of the notes.
+
+\begin{figure*}[htp]
+    \centering
+    \includegraphics[width=\textwidth]{deconvolution.png}
+    \caption{Deconvolution of the (power) CQT-spectrogram of 12 acoustic bass notes playing from C1 to B1, into a pitch-independent envelope and an energy-normalized pitch component.}
+    \label{fig:deconvolution}
+\end{figure*}
+
 This deconvolution process can also be thought of the normalization of the log-spectrum by the magnitude of its FT (which here would correspond to the FT of the envelope) leading to a sharper log-spectrum (which here would correspond to the pitch component), in the manner of the generalized cross-correlation phase transform (GCC-PHAT) method which aims at normalizing a cross-correlation function by its magnitude spectrum to sharpen the cross-correlation peaks \cite{knapp1976}.
@@
 
-Once the CQT spectrum has been decomposed into a spectral envelope and a pitch component, we can then assumed that most of the pitch information has been removed from the envelope which now mostly contains the timbre information.
-We can think of the spectral envelope as a pitch-normalized spectrum ...
+%HERE!!!
+Once the CQT-spectrum has been decomposed into a spectral envelope and a pitch component, we can then assumed that most of the pitch information has been removed from the envelope which now mostly contains the timbre information.
+We can think of the spectral envelope as a pitch-normalized log-spectrum where the frequency components have been shifted down to the origin of the spectrum.
+
+Equation \ref{eq:extraction} $O_r$ the octave resolution, or the number of frequency channels per octave (typically, 12, 24, 36, etc.), $N_c$ the number of coefficients
+
+\begin{equation}
+\label{eq:extraction}
+i = O_r \log_2(k) \text{ for } 1 \le k \le N_c
+\end{equation}
 

### 2021-06-23 IEEE SP Magazine/06-2021 - Draft/article.tex
Almost done with the extraction subsection.

@@
 
-This convolution process can be thought of a source-filter model \cite{fant1970} which is here not applied in the time domain but in the frequency domain, with the source and the filter being the pitch component and the envelope, respectively.
+This convolution process can be thought of as a source-filter model \cite{fant1970} which is here not applied in the time domain but in the frequency domain, with the source and the filter being the pitch component and the envelope, respectively.
 
@@
 \begin{split}
-\begin{cases}
+& \begin{cases}
 X = E * P \\
 X' = E' * P' \\
-\end{cases}
-\Rightarrow E \approx E'
+\end{cases} \\
+& \Rightarrow E \approx E'
 \end{split}
@@
 \begin{split}
-\begin{cases}
+& \begin{cases}
 \mathcal{F}(X) = \mathcal{F}(E) \cdot \mathcal{F}(P) \\
 \mathcal{F}(X') = \mathcal{F}(E') \cdot \mathcal{F}(P') \\
-\end{cases}
-\Rightarrow \mathcal{F}(E) \approx \mathcal{F}(E')
+\end{cases} \\
+& \Rightarrow \mathcal{F}(E) \approx \mathcal{F}(E')
 \end{split}
@@
 \begin{split}
-\begin{cases}
+& \begin{cases}
 \mathcal{F}(X) = |\mathcal{F}(X)| \cdot e^{j Arg(\mathcal{F}(X))} \\
 \mathcal{F}(X') = |\mathcal{F}(X')| \cdot e^{j Arg(\mathcal{F}(X'))} \\
-\end{cases}
-\Rightarrow |\mathcal{F}(X)| \approx |\mathcal{F}(X')|
+\end{cases} \\
+& \Rightarrow |\mathcal{F}(X)| \approx |\mathcal{F}(X')|
 \end{split}
@@
 \begin{split}
-\Rightarrow 
+& \Rightarrow 
 \begin{cases}
-\mathcal{F}(E) \approx |\mathcal{F}(X)| &\Rightarrow E \approx \mathcal{F}^{-1}(|\mathcal{F}(X)|) \\
-\mathcal{F}(P) \approx e^{j Arg(\mathcal{F}(X))} &\Rightarrow P \approx \mathcal{F}^{-1}(e^{j Arg(\mathcal{F}(X))}) \\
+\mathcal{F}(E) \approx |\mathcal{F}(X)| \\
+\mathcal{F}(P) \approx e^{j Arg(\mathcal{F}(X))} \\
+\end{cases} \\
+& \Rightarrow
+\begin{cases}
+E \approx \mathcal{F}^{-1}(|\mathcal{F}(X)|) \\
+P \approx \mathcal{F}^{-1}(e^{j Arg(\mathcal{F}(X))}) \\
 \end{cases}
@@
 
-Figure \ref{fig:deconvolution} shows and example of deconvolution of the (power) CQT-spectrogram into its envelope and pitch component.
-The CQT-spectogram was computed from a 48-second audio signal created by concatenating 12 4-second acoustic bass notes, playing successively from C1 (32.70 Hz) to B1 (61.74 Hz).
-The notes come from the NSynth dataset \cite{engel2017}, for instrument \texttt{bass\_acoustic\_000}, MIDI numbers \texttt{024} to \texttt{035}, and velocity number \texttt{075}.
-As we can see, the envelope gets pitch-normalized, shifting all the frequency components to the origin of the log-spectrum (essentially bringing all the notes to C1), while the pitch component gets energy-normalized, leaving mostly the fundamental frequency of the notes.
+Figure \ref{fig:deconvolution} shows an example of deconvolution of a (power) CQT-spectrogram into its envelope and pitch component.
+The CQT-spectogram was computed from a 48-second audio signal created by concatenating 12 4-second notes from an acoustic bass playing from C1 (32.70 Hz) to B1 (61.74 Hz) in ascending order.
+The notes come from the NSynth dataset \cite{engel2017} and correspond to instrument id \texttt{bass\_acoustic\_000}, MIDI numbers \texttt{024} to \texttt{035}, and velocity number \texttt{075}.
+As we can see, the envelope looks as if the CQT-spectrogram has been normalized in pitch, with all the notes being brought down to the lowest frequency, here C1, while the pitch component looks as if the CQT-spectrogram has been stripped down from all its energy, leaving mostly the fundamental frequencies of the notes.
 
@@
     \includegraphics[width=\textwidth]{deconvolution.png}
-    \caption{Deconvolution of the (power) CQT-spectrogram of 12 acoustic bass notes playing from C1 to B1, into a pitch-independent envelope and an energy-normalized pitch component.}
+    \caption{Deconvolution of the CQT-spectrogram (left plot, shown in dB) of 12 acoustic bass notes playing from C1 to B1, into a pitch-independent envelope (middle plot, shown in dB) and an energy-normalized pitch component (right plot, shown in $[0, 1]$).}
     \label{fig:deconvolution}
@@
 
-This deconvolution process can also be thought of the normalization of the log-spectrum by the magnitude of its FT (which here would correspond to the FT of the envelope) leading to a sharper log-spectrum (which here would correspond to the pitch component), in the manner of the generalized cross-correlation phase transform (GCC-PHAT) method which aims at normalizing a cross-correlation function by its magnitude spectrum to sharpen the cross-correlation peaks \cite{knapp1976}.
+This deconvolution process can also be thought of as the normalization of the log-spectrum by the magnitude of its FT (which here would correspond to the FT of the envelope) leading to a sharper log-spectrum (which here would correspond to the pitch component), in the manner of the generalized cross-correlation phase transform (GCC-PHAT) method which aims at normalizing a cross-correlation function by its magnitude spectrum to sharpen the cross-correlation peaks \cite{knapp1976}.
 
@@
 
-%HERE!!!
-Once the CQT-spectrum has been decomposed into a spectral envelope and a pitch component, we can then assumed that most of the pitch information has been removed from the envelope which now mostly contains the timbre information.
-We can think of the spectral envelope as a pitch-normalized log-spectrum where the frequency components have been shifted down to the origin of the spectrum.
-
-Equation \ref{eq:extraction} $O_r$ the octave resolution, or the number of frequency channels per octave (typically, 12, 24, 36, etc.), $N_c$ the number of coefficients
-
+Once the CQT-spectrum has been decomposed into an envelope and a pitch component, we can then assume that most of the pitch information has been removed from the envelope which now mostly contains the timbre information.
+As shown in the middle plot of Figure \ref{fig:deconvolution}, the envelope can be thought of as a pitch-normalized log-spectrum where the spectral components have been shifted down to the origin of the spectrum.
+If we know what octave resolution was used when computing the CQT-spectrum, i.e., the number of bins per octave, we can easily infer the locations of the harmonics in the envelope, which should then judiciously characterize the timbre of the musical instrument in question.
+Hence, the CQT-SECs.
+Equation \ref{eq:extraction} shows how to derive the indices of the harmonics given $O_r$, the octave resolution, and $N_c$, the number of desired coefficients, and consequently, the CQT-SECs from the envelope $E$.
 \begin{equation}
 \label{eq:extraction}
-i = O_r \log_2(k) \text{ for } 1 \le k \le N_c
+\begin{split}
+\begin{cases}
+i = O_r \log_2(k) \\
+\text{CQT-SEC}_k = E(i) \\
+\end{cases}
+\quad 1 \le k \le N_c
+\end{split}
 \end{equation}

### 2021-06-23 IEEE SP Magazine/06-2021 - Draft/article.tex
Still working on extraction subsection.

@@
 It is generally associated with the spectral (but also temporal) envelope and is typically assumed to be independent from the pitch (but also the loudness) of the sound \cite{moore2004}.
-In this article, we will show how to design a simple but functional pitch-independent timbre feature which is well-adapted to musical data, by deriving it from the constant-Q transform (CQT) \cite{brown1991, brown1992}, a log-frequency transform which matches the Western music scale.
+In this article, we will show how to design a simple but functional pitch-independent timbre feature which is well-adapted to musical data, by deriving it from the constant-Q transform (CQT) \cite{brown1991, brown1992}, a log-frequency transform which aligns with the Western music scale.
 We will show how to decompose the CQT-spectrum into an energy-normalized pitch component and a pitch-independent spectral envelope, the latter from which we will extract a number of timbral coefficients.
@@
 & \begin{cases}
-\mathcal{F}(X) = \mathcal{F}(E) \cdot \mathcal{F}(P) \\
-\mathcal{F}(X') = \mathcal{F}(E') \cdot \mathcal{F}(P') \\
+\mathcal{F}(X) = \mathcal{F}(E * P) = \mathcal{F}(E) \cdot \mathcal{F}(P) \\
+\mathcal{F}(X') = \mathcal{F}(E' * P') = \mathcal{F}(E') \cdot \mathcal{F}(P') \\
 \end{cases} \\
@@
 This implies that the magnitude of the FT of the log-spectrum for a musical object and for a pitch-shifted version of it would be equal.
-This is summarized in Equation \ref{eq:observation3}, where $|.|$ and $Arg(.)$ represent the modulus and argument, respectively, for a complex array.
+This is summarized in Equation \ref{eq:observation3}, where $|.|$ and $Arg(.)$ represent the modulus and argument, respectively, for a complex array, and $j$, the imaginary unit.
 \begin{equation}
@@
 
-Figure \ref{fig:deconvolution} shows an example of deconvolution of a (power) CQT-spectrogram into its envelope and pitch component.
-The CQT-spectogram was computed from a 48-second audio signal created by concatenating 12 4-second notes from an acoustic bass playing from C1 (32.70 Hz) to B1 (61.74 Hz) in ascending order.
+Figure \ref{fig:deconvolution} shows an example of deconvolution of a CQT-spectrogram into its envelope and pitch component.
+The CQT-spectogram was computed from an audio signal created by concatenating 12 4-second notes from an acoustic bass playing from C1 (32.70 Hz) to B1 (61.74 Hz) in ascending order.
 The notes come from the NSynth dataset \cite{engel2017} and correspond to instrument id \texttt{bass\_acoustic\_000}, MIDI numbers \texttt{024} to \texttt{035}, and velocity number \texttt{075}.
 As we can see, the envelope looks as if the CQT-spectrogram has been normalized in pitch, with all the notes being brought down to the lowest frequency, here C1, while the pitch component looks as if the CQT-spectrogram has been stripped down from all its energy, leaving mostly the fundamental frequencies of the notes.
+Note that in practice, we use a power CQT-spectrogram and we take the real part of the envelope and pitch component to ensure real values.
 
@@
 Once the CQT-spectrum has been decomposed into an envelope and a pitch component, we can then assume that most of the pitch information has been removed from the envelope which now mostly contains the timbre information.
-As shown in the middle plot of Figure \ref{fig:deconvolution}, the envelope can be thought of as a pitch-normalized log-spectrum where the spectral components have been shifted down to the origin of the spectrum.
-If we know what octave resolution was used when computing the CQT-spectrum, i.e., the number of bins per octave, we can easily infer the locations of the harmonics in the envelope, which should then judiciously characterize the timbre of the musical instrument in question.
-Hence, the CQT-SECs.
-Equation \ref{eq:extraction} shows how to derive the indices of the harmonics given $O_r$, the octave resolution, and $N_c$, the number of desired coefficients, and consequently, the CQT-SECs from the envelope $E$.
+As shown in the middle plot of Figure \ref{fig:deconvolution}, the envelope can be thought of as a pitch-normalized log-spectrum where the spectral components, in particular the harmonics which contain most of the energy from the musical instrument, have been essentially brought down to the same note level.
+Given the octave resolution that was used to compute the CQT-spectrum, i.e., the number of bins per octave, we can then easily infer the locations of those harmonics in the envelope.
+We can subsequently extract these harmonics, or \textit{spectral envelope coefficients}, from the envelope, and therefore obtain a judicious compact feature for characterizing the timbre of the musical instrument.
+Equation \ref{eq:extraction} shows how to derive the indices of the spectral envelope coefficients given $O_r$, the octave resolution, and $N_c$, the number of desired coefficients, and consequently extract the CQT-SECs from the envelope $E$, with $\log_2(.)$ and $round(.)$ representing the binary logarithm and the round function, respectively.
 \begin{equation}
@@
 \begin{cases}
-i = O_r \log_2(k) \\
+i = round(O_r \log_2(k)) \\
 \text{CQT-SEC}_k = E(i) \\
@@
 
-Briefly explain what how the MFCCs are computed \cite{mermelstein1976}
+Figure \ref{fig:extraction} shows an example of CQT-SECs ...
+
+in dB
+
+\begin{figure*}[htp]
+    \centering
+    \includegraphics[width=\textwidth]{extraction.png}
+    \caption{Example of CQT-SECs ... (left plot, shown in dB), and MFCCs ... (right plot).}
+    \label{fig:extraction}
+\end{figure*}
+
+We recall how the MFCCs are computed ...\cite{mermelstein1976}
 mel scale: \cite{stevens1937}

### 2021-06-29 IEEE SP Magazine/06-2021 - Draft/article.tex
Worked more on the solution section.

@@
 It is generally associated with the spectral (but also temporal) envelope and is typically assumed to be independent from the pitch (but also the loudness) of the sound \cite{moore2004}.
-In this article, we will show how to design a simple but functional pitch-independent timbre feature which is well-adapted to musical data, by deriving it from the constant-Q transform (CQT) \cite{brown1991, brown1992}, a log-frequency transform which aligns with the Western music scale.
+In this article, we will show how to design a simple but functional pitch-independent timbre feature which is well-adapted to musical data, by deriving it from the constant-Q transform (CQT) \cite{brown1991, brown1992}, a log-frequency transform which matches the equal temperament musical scale.
 We will show how to decompose the CQT-spectrum into an energy-normalized pitch component and a pitch-independent spectral envelope, the latter from which we will extract a number of timbral coefficients.
@@
 In particular, they are derived through an old process which makes use of the mel scale, a perceptual scale experimentally designed 85 years ago to approximate the human auditory system's response \cite{stevens1937}.
-More recently, a number of data-driven approaches attempted to learn some timbral representations from musical data, but generally in terms of implicit embeddings which are tied to a specific trained model \cite{engel2017}, \cite{pons2017}, and not necessarily as explicit and interpretable features such as the MFCCs, which are still usually preferred as the go-to feature to characterize musical timbre by MIR practitioners.
+More recently, a number of data-driven approaches attempted to learn some timbral representations from musical data, but generally in terms of implicit embeddings which are tied to a specific trained model \cite{engel2017}, \cite{pons2017}, and not necessarily as explicit and interpretable features such as the MFCCs, which are still usually preferred as the go-to feature to characterize timbre by MIR practitioners.
 
@@
 We propose here the CQT-SECs, a novel timbre feature which is well-adapted to musical data, pitch-independent, simple to compute, interpretable, and functional.
-We will show how to derive it from the CQT, a frequency transform with a logarithmic resolution which matches the notes of Western music scale \cite{brown1991, brown1992}, by first decomposing the CQT log-spectrum into a pitch-independent spectral envelope and an energy-normalized pitch component, and then extracting a number of timbral coefficients from the spectral envelope.
+We will show how to derive it from the CQT, a frequency transform with a logarithmic resolution which matches the notes of the equal temperament, a musical scale typically used in Western music \cite{brown1991, brown1992}.
+We will first show how to decompose the CQT-spectrum into a pitch-independent spectral envelope and an energy-normalized pitch component, and then extract a number of timbral coefficients from the spectral envelope.
 
@@
 Figure \ref{fig:deconvolution} shows an example of deconvolution of a CQT-spectrogram into its envelope and pitch component.
-The CQT-spectogram was computed from an audio signal created by concatenating 12 4-second notes from an acoustic bass playing from C1 (32.70 Hz) to B1 (61.74 Hz) in ascending order.
+The CQT-spectogram was computed from an audio signal created by concatenating 12 4-second notes of an acoustic bass playing from C1 (32.70 Hz) to B1 (61.74 Hz) in ascending order.
 The notes come from the NSynth dataset \cite{engel2017} and correspond to instrument id \texttt{bass\_acoustic\_000}, MIDI numbers \texttt{024} to \texttt{035}, and velocity number \texttt{075}.
-As we can see, the envelope looks as if the CQT-spectrogram has been normalized in pitch, with all the notes being brought down to the lowest frequency, here C1, while the pitch component looks as if the CQT-spectrogram has been stripped down from all its energy, leaving mostly the fundamental frequencies of the notes.
-Note that in practice, we use a power CQT-spectrogram and we take the real part of the envelope and pitch component to ensure real values.
+As we can see, the envelope looks as if the CQT-spectrogram has been normalized in pitch, with all the notes being brought down to the lowest frequency (here, corresponding to C1); while the pitch component looks as if the CQT-spectrogram has been stripped down from all its energy, leaving mostly the fundamental frequencies of the notes.
+Note that, in practice, we use a power CQT-spectrogram (magnitude to the power of 2) and take the real part of the envelope and pitch component to ensure real values.
+We also note that this deconvolution can potentially be further refined, for example, by zeroing the few negative values in the pitch component to only have values in $[0, 1]$, and using it to re-derive the spectral envelope from the log-spectrum.
 
@@
     \includegraphics[width=\textwidth]{deconvolution.png}
-    \caption{Deconvolution of the CQT-spectrogram (left plot, shown in dB) of 12 acoustic bass notes playing from C1 to B1, into a pitch-independent envelope (middle plot, shown in dB) and an energy-normalized pitch component (right plot, shown in $[0, 1]$).}
+    \caption{Deconvolution of the CQT-spectrogram (left plot, shown in dB) of 12 acoustic bass notes playing from C1 to B1, into a pitch-independent spectral envelope (middle plot, shown in dB) and an energy-normalized pitch component (right plot, shown in $[0, 1]$).}
     \label{fig:deconvolution}
@@
 Once the CQT-spectrum has been decomposed into an envelope and a pitch component, we can then assume that most of the pitch information has been removed from the envelope which now mostly contains the timbre information.
-As shown in the middle plot of Figure \ref{fig:deconvolution}, the envelope can be thought of as a pitch-normalized log-spectrum where the spectral components, in particular the harmonics which contain most of the energy from the musical instrument, have been essentially brought down to the same note level.
+As shown in the middle plot of Figure \ref{fig:deconvolution}, the envelope can be thought of as a pitch-normalized log-spectrum where the spectral components, in particular, the harmonics which contain most of the energy of the musical instrument, have been essentially brought down to the same note level.
 Given the octave resolution that was used to compute the CQT-spectrum, i.e., the number of bins per octave, we can then easily infer the locations of those harmonics in the envelope.
-We can subsequently extract these harmonics, or \textit{spectral envelope coefficients}, from the envelope, and therefore obtain a judicious compact feature for characterizing the timbre of the musical instrument.
+We can subsequently extract these harmonics, or \textit{spectral envelope coefficients}, from the envelope, and therefore obtain a compact and meaningful feature for characterizing the timbre of the musical instrument.
 Equation \ref{eq:extraction} shows how to derive the indices of the spectral envelope coefficients given $O_r$, the octave resolution, and $N_c$, the number of desired coefficients, and consequently extract the CQT-SECs from the envelope $E$, with $\log_2(.)$ and $round(.)$ representing the binary logarithm and the round function, respectively.

### 2021-06-30 IEEE SP Magazine/06-2021 - Draft/article.tex
Updated article.

@@
 \IEEEPARstart{T}{imbre} is the attribute of sound which makes, for example, two musical instruments playing the same note sound different.
-It is generally associated with the spectral (but also temporal) envelope and is typically assumed to be independent from the pitch (but also the loudness) of the sound \cite{moore2004}.
-In this article, we will show how to design a simple but functional pitch-independent timbre feature which is well-adapted to musical data, by deriving it from the constant-Q transform (CQT) \cite{brown1991, brown1992}, a log-frequency transform which matches the equal temperament musical scale.
-We will show how to decompose the CQT-spectrum into an energy-normalized pitch component and a pitch-independent spectral envelope, the latter from which we will extract a number of timbral coefficients.
-We will then evaluate the discriminative power of these CQT-spectral envelope coefficients (CQT-SEC) on the NSynth dataset \cite{engel2017}, a large-scale dataset of musical notes which is publicly available, comparing them with the mel-frequency cepstral coefficients (MFCCs) \cite{mermelstein1976}, features originally designed for speech recognition but commonly used to characterize timbre in music.
+It is typically associated with the spectral (but also the temporal) envelope and assumed to be independent from the pitch (but also the loudness) of the sound \cite{moore2004}.
+In this article, we will show how to design a simple but functional pitch-independent timbre feature which is well adapted to musical data, by deriving it from the constant-Q transform (CQT), a log-frequency transform which matches the equal tempered musical scale \cite{brown1991, brown1992}.
+We will show how to decompose the CQT spectrum into an energy-normalized pitch component and a pitch-independent spectral envelope, the latter from which we will extract a number of timbral coefficients.
+We will then evaluate the discriminative power of these \textit{CQT spectral envelope coefficients (CQT-SEC)} on the NSynth dataset \cite{engel2017}, a large-scale dataset of musical notes which is publicly available, comparing them with the mel-frequency cepstral coefficients (MFCCs) \cite{mermelstein1976}, features originally designed for speech recognition but commonly used to characterize timbre in music.
 
@@
 
-A timbre feature which is well-adapted to musical data, pitch-independent, and with high discriminative power can find uses in a number of applications, such as similarity detection, sound recognition, and audio classification, in particular, of musical instruments.
-Additionally, the ability to decompose the spectrum of a sound (here, the CQT-spectrum) into a pitch-independent spectral envelope and an energy-normalized pitch component can be useful for analysis, transformation, and resynthesis of music signals.
+A timbre feature which is well adapted to musical data, pitch-independent, and with high discriminative power can find uses in a number of applications, such as similarity detection, sound recognition, and audio classification, in particular, of musical instruments.
+Additionally, the ability to decompose the spectrum of a sound (here, the CQT spectrum) into a pitch-independent spectral envelope and an energy-normalized pitch component can be useful for analysis, transformation, and resynthesis of music signals.
 The energy-normalized pitch component can also potentially be used for tasks such as pitch identification, melody extraction, and chord recognition.
@@
 Researchers in MIR proposed a number of descriptors to characterize one or more aspects of timbre \cite{peeters2011}, but they mostly resort to using the MFCCs when they need one simple timbre feature \cite{mueller2007}.
-While the MFCCs were shown to be helpful in some MIR tasks, they were initially designed for speech processing applications \cite{mermelstein1976} and are not necessarily well-adapted to musical data.
+While the MFCCs were shown to be practical in some MIR tasks, they were initially designed for speech processing applications \cite{mermelstein1976} and are not necessarily well adapted to musical data.
 In particular, they are derived through an old process which makes use of the mel scale, a perceptual scale experimentally designed 85 years ago to approximate the human auditory system's response \cite{stevens1937}.
-More recently, a number of data-driven approaches attempted to learn some timbral representations from musical data, but generally in terms of implicit embeddings which are tied to a specific trained model \cite{engel2017}, \cite{pons2017}, and not necessarily as explicit and interpretable features such as the MFCCs, which are still usually preferred as the go-to feature to characterize timbre by MIR practitioners.
+More recently, a number of data-driven approaches attempted to learn some timbral representations from musical data, but typically in terms of implicit embeddings which are tied to a specific trained model \cite{engel2017}, \cite{pons2017}, and not necessarily as explicit and interpretable features such as the MFCCs, which are still usually preferred as the go-to feature to characterize timbre by MIR practitioners.
 
@@
 
-We propose here the CQT-SECs, a novel timbre feature which is well-adapted to musical data, pitch-independent, simple to compute, interpretable, and functional.
+We propose the CQT-SECs, a novel timbre feature which is well adapted to musical data, pitch-independent, simple to compute, interpretable, and functional.
 We will show how to derive it from the CQT, a frequency transform with a logarithmic resolution which matches the notes of the equal temperament, a musical scale typically used in Western music \cite{brown1991, brown1992}.
-We will first show how to decompose the CQT-spectrum into a pitch-independent spectral envelope and an energy-normalized pitch component, and then extract a number of timbral coefficients from the spectral envelope.
+We will first show how to decompose the CQT spectrum into a pitch-independent spectral envelope and an energy-normalized pitch component, and then extract a number of timbral coefficients from the spectral envelope.
 
@@
 
-We start with the assumption that a log-spectrum $X$, in particular, the CQT-spectrum, can be represented as the convolution between a pitch-independent spectral envelope $E$ (which mostly contains the timbre information) and an energy-normalized pitch component $P$ (which mostly contains the pitch information), as shown in Equation \ref{eq:assumption}.
+We start with the assumption that a log spectrum $X$, in particular, the CQT spectrum, can be represented as the convolution between a pitch-independent spectral envelope $E$ (which mostly contains the timbre information) and an energy-normalized pitch component $P$ (which mostly contains the pitch information), as shown in Equation \ref{eq:assumption}, where $*$ represents the convolution operation.
 \begin{equation}
@@
 
-\emph{Observation 1: A pitch change in the audio translates to a linear shift in the log-spectrum \cite{brown1991, brown1992}.}
+\emph{Observation 1: A pitch change in the audio translates to a linear shift in the log spectrum \cite{brown1991, brown1992}.}
 
 Assuming that pitch and timbre are independent, this implies that the same musical object at different pitches would have a similar envelope but a shifted pitch component (while two different musical objects at the same pitch would have different envelopes but a similar pitch component).
-This is summarized in Equation \ref{eq:observation1}, where $X$, $E$, $P$ and $X'$, $E'$, $P'$ represent the log-spectrum, envelope, and pitch component for a musical object and for a pitch-shifted version of the same musical object, respectively.
+This is summarized in Equation \ref{eq:observation1}, where $X$, $E$, $P$ and $X'$, $E'$, $P'$ represent the log spectrum, envelope, and pitch component for a musical object and for a pitch-shifted version of the same musical object, respectively.
 \begin{equation}
@@
 
-\emph{Observation 2: the FT of the convolution of two functions is equal to the point-wise product of the FTs of the two functions, a property also known as the convolution theorem \cite{proakis1995}.}
+\emph{Observation 2: the FT of the convolution between two functions is equal to the pointwise product between the FTs of the two functions, a property also known as the convolution theorem \cite{proakis1995}.}
 
-This implies that the FT of the log-spectrum is equal to the point-wise product of the FT of the envelope and the FT of the pitch component.
+This implies that the FT of the log spectrum is equal to the pointwise product between the FT of the envelope and the FT of the pitch component.
 Given the first observation, this further implies that the FT of the envelope for a musical object and for a pitch-shifted version of it would be equal.
-This is summarized in Equation \ref{eq:observation2}, where $\mathcal{F}(.)$ represents the FT function.
+This is summarized in Equation \ref{eq:observation2}, where $\mathcal{F}(.)$ represents the FT function and $\cdot$ the pointwise product.
 \begin{equation}
@@
 
-This implies that the magnitude of the FT of the log-spectrum for a musical object and for a pitch-shifted version of it would be equal.
+This implies that the magnitude of the FT of the log spectrum for a musical object and for a pitch-shifted version of it would be equal.
 This is summarized in Equation \ref{eq:observation3}, where $|.|$ and $Arg(.)$ represent the modulus and argument, respectively, for a complex array, and $j$, the imaginary unit.
@@
 
-Given the previous observations, we can therefore conclude that the FT of the envelope could be approximated by the magnitude of the FT of the log-spectrum, while the FT of the pitch component could be approximated by the phase component.
+Given the previous observations, we can therefore conclude that the FT of the envelope could be approximated by the magnitude of the FT of the log spectrum, while the FT of the pitch component could be approximated by the phase component.
 This finally gives us the estimates for the envelope and the pitch component, after taking their inverse FTs, as shown in Equation \ref{eq:conclusion}, where $\mathcal{F}^{-1}(.)$ represents the inverse FT function.
@@
 
-Figure \ref{fig:deconvolution} shows an example of deconvolution of a CQT-spectrogram into its envelope and pitch component.
-The CQT-spectogram was computed from an audio signal created by concatenating 12 4-second notes of an acoustic bass playing from C1 (32.70 Hz) to B1 (61.74 Hz) in ascending order.
-The notes come from the NSynth dataset \cite{engel2017} and correspond to instrument id \texttt{bass\_acoustic\_000}, MIDI numbers \texttt{024} to \texttt{035}, and velocity number \texttt{075}.
-As we can see, the envelope looks as if the CQT-spectrogram has been normalized in pitch, with all the notes being brought down to the lowest frequency (here, corresponding to C1); while the pitch component looks as if the CQT-spectrogram has been stripped down from all its energy, leaving mostly the fundamental frequencies of the notes.
-Note that, in practice, we use a power CQT-spectrogram (magnitude to the power of 2) and take the real part of the envelope and pitch component to ensure real values.
-We also note that this deconvolution can potentially be further refined, for example, by zeroing the few negative values in the pitch component to only have values in $[0, 1]$, and using it to re-derive the spectral envelope from the log-spectrum.
+Figure \ref{fig:deconvolution} shows an example of deconvolution of a CQT spectrogram into its envelope and pitch component.
+The CQT spectogram was computed from an audio signal created by concatenating 12 4-second notes of an acoustic bass playing from C1 (32.70 Hz) to B1 (61.74 Hz) in ascending order.
+The CQT spectrogram was computed using librosa \cite{mcfee2015}
+
+HERE!!!
+
+and the notes come from the NSynth dataset \cite{engel2017} and correspond to instrument id \texttt{bass\_acoustic\_000}, MIDI numbers \texttt{024} to \texttt{035}, and velocity number \texttt{075}.
+As we can see, the envelope looks as if the CQT spectrogram has been normalized in pitch, with all the notes being brought down to the lowest frequency (here, corresponding to C1); while the pitch component looks as if the CQT spectrogram has been stripped down from all its energy, leaving mostly the fundamental frequencies of the notes.
+Note that, in practice, we use a power CQT spectrogram (magnitude to the power of 2) and take the real part of the envelope and pitch component to ensure real values.
+We also note that this deconvolution can potentially be further refined, for example, by zeroing the few negative values in the pitch component to only have values in $[0, 1]$, and using it to re-derive the spectral envelope from the log spectrum.
 
@@
     \includegraphics[width=\textwidth]{deconvolution.png}
-    \caption{Deconvolution of the CQT-spectrogram (left plot, shown in dB) of 12 acoustic bass notes playing from C1 to B1, into a pitch-independent spectral envelope (middle plot, shown in dB) and an energy-normalized pitch component (right plot, shown in $[0, 1]$).}
+    \caption{Deconvolution of the CQT spectrogram (left plot, shown in dB) of 12 acoustic bass notes playing from C1 to B1, into a pitch-independent spectral envelope (middle plot, shown in dB) and an energy-normalized pitch component (right plot, shown in $[0, 1]$).}
     \label{fig:deconvolution}
@@
 
-This deconvolution process can also be thought of as the normalization of the log-spectrum by the magnitude of its FT (which here would correspond to the FT of the envelope) leading to a sharper log-spectrum (which here would correspond to the pitch component), in the manner of the generalized cross-correlation phase transform (GCC-PHAT) method which aims at normalizing a cross-correlation function by its magnitude spectrum to sharpen the cross-correlation peaks \cite{knapp1976}.
+This deconvolution process can also be thought of as the normalization of the log spectrum by the magnitude of its FT (which here would correspond to the FT of the envelope) leading to a sharper log spectrum (which here would correspond to the pitch component), in the manner of the generalized cross-correlation phase transform (GCC-PHAT) method which aims at normalizing a cross-correlation function by its magnitude spectrum to sharpen the cross-correlation peaks \cite{knapp1976}.
 
@@
 
-Once the CQT-spectrum has been decomposed into an envelope and a pitch component, we can then assume that most of the pitch information has been removed from the envelope which now mostly contains the timbre information.
-As shown in the middle plot of Figure \ref{fig:deconvolution}, the envelope can be thought of as a pitch-normalized log-spectrum where the spectral components, in particular, the harmonics which contain most of the energy of the musical instrument, have been essentially brought down to the same note level.
-Given the octave resolution that was used to compute the CQT-spectrum, i.e., the number of bins per octave, we can then easily infer the locations of those harmonics in the envelope.
-We can subsequently extract these harmonics, or \textit{spectral envelope coefficients}, from the envelope, and therefore obtain a compact and meaningful feature for characterizing the timbre of the musical instrument.
+The envelope resulting from the deconvolution of the CQT spectrum can be thought of as a pitch-normalized CQT spectrum, where the spectral components, in particular, the harmonics which contain most of the energy of a musical instrument, have been essentially brought down to the same lowest note level.
+Given the octave resolution that was used to compute the CQT spectrum, i.e., the number of bins per octave, we can then easily infer the locations of those harmonics in the envelope \cite{brown1991, brown1992}.
+We can subsequently extract these harmonics, or \textit{spectral envelope coefficients}, from the envelope, and therefore obtain a compact and interpretable feature for characterizing the timbre of the musical instrument.
 Equation \ref{eq:extraction} shows how to derive the indices of the spectral envelope coefficients given $O_r$, the octave resolution, and $N_c$, the number of desired coefficients, and consequently extract the CQT-SECs from the envelope $E$, with $\log_2(.)$ and $round(.)$ representing the binary logarithm and the round function, respectively.
@@
 
-Figure \ref{fig:extraction} shows an example of CQT-SECs ...
-
-in dB
+Figure \ref{fig:extraction} shows an example of CQT-SECs, on the left plot.
+20 coefficients were extracted from the spectral envelope obtained following the deconvolution of the CQT spectrogram of the musical signal shown in Figure \ref{fig:deconvolution}.
+The coefficients are shown in dB to be comparable to the envelope which was also shown in dB.
+These coefficients essentially correspond to the harmonics of the musical instrument which contain most of its spectral energy and can therefore be a good representation of its timbre.
+For comparison, we also show the MFCCs computed from the same musical signal, on the right plot.
+20 coefficients were computed using
 
@@
     \includegraphics[width=\textwidth]{extraction.png}
-    \caption{Example of CQT-SECs ... (left plot, shown in dB), and MFCCs ... (right plot).}
+    \caption{CQT-SECs extracted from the envelope obtained following the deconvolution shown in Figure \ref{fig:deconvolution} (left plot, shown in dB) and MFCCs computed from the same musical signal (right plot).}
     \label{fig:extraction}
@@
 mel scale: \cite{stevens1937}
+
+using librosa \cite{mcfee2015}
 

### 2021-06-30 IEEE SP Magazine/06-2021 - Draft/article.tex
Finished Solution section.

@@
 The CQT spectogram was computed from an audio signal created by concatenating 12 4-second notes of an acoustic bass playing from C1 (32.70 Hz) to B1 (61.74 Hz) in ascending order.
-The CQT spectrogram was computed using librosa \cite{mcfee2015}
-
-HERE!!!
-
-and the notes come from the NSynth dataset \cite{engel2017} and correspond to instrument id \texttt{bass\_acoustic\_000}, MIDI numbers \texttt{024} to \texttt{035}, and velocity number \texttt{075}.
-As we can see, the envelope looks as if the CQT spectrogram has been normalized in pitch, with all the notes being brought down to the lowest frequency (here, corresponding to C1); while the pitch component looks as if the CQT spectrogram has been stripped down from all its energy, leaving mostly the fundamental frequencies of the notes.
+The notes come from the NSynth dataset \cite{engel2017} and correspond to instrument id \texttt{bass\_acoustic\_000}, MIDI numbers \texttt{024} to \texttt{035}, and velocity number \texttt{075}.
+The CQT spectrogram was computed using librosa \cite{mcfee2015,schoerkhuber2010}, with a sampling rate of 16 kHz, a hop length of 512 samples, a minimum frequency of 32.70 Hz (corresponding to C1), 95 frequency bins, and 12 bins per octave.
+As we can see, the envelope looks as if the CQT spectrogram has been normalized in pitch, with all the notes being brought down to the lowest frequency (corresponding to C1); while the pitch component looks as if the CQT spectrogram has been stripped down from all its energy, leaving mostly the fundamental frequencies of the notes.
 Note that, in practice, we use a power CQT spectrogram (magnitude to the power of 2) and take the real part of the envelope and pitch component to ensure real values.
-We also note that this deconvolution can potentially be further refined, for example, by zeroing the few negative values in the pitch component to only have values in $[0, 1]$, and using it to re-derive the spectral envelope from the log spectrum.
+We also note that this deconvolution can potentially be further refined, for example, by zeroing the few negative values in the pitch component and using it to re-derive the spectral envelope from the log spectrum.
 
@@
 
-The envelope resulting from the deconvolution of the CQT spectrum can be thought of as a pitch-normalized CQT spectrum, where the spectral components, in particular, the harmonics which contain most of the energy of a musical instrument, have been essentially brought down to the same lowest note level.
-Given the octave resolution that was used to compute the CQT spectrum, i.e., the number of bins per octave, we can then easily infer the locations of those harmonics in the envelope \cite{brown1991, brown1992}.
-We can subsequently extract these harmonics, or \textit{spectral envelope coefficients}, from the envelope, and therefore obtain a compact and interpretable feature for characterizing the timbre of the musical instrument.
+The spectral envelope resulting from the deconvolution of the CQT spectrum can be thought of as a pitch-normalized CQT spectrum where the spectral content, in particular, the harmonics which represent most of the energy of the musical instrument, has been essentially brought down to the same lowest note level.
+Given the octave resolution which was used when computing the CQT spectrum, i.e., the number of bins per octave, we can then easily infer the locations of those harmonics in the envelope \cite{brown1991, brown1992}.
+We can subsequently extract these harmonics or \textit{spectral envelope coefficients} from the envelope and therefore obtain a compact and interpretable feature for characterizing the timbre of the musical instrument.
 Equation \ref{eq:extraction} shows how to derive the indices of the spectral envelope coefficients given $O_r$, the octave resolution, and $N_c$, the number of desired coefficients, and consequently extract the CQT-SECs from the envelope $E$, with $\log_2(.)$ and $round(.)$ representing the binary logarithm and the round function, respectively.
@@
 Figure \ref{fig:extraction} shows an example of CQT-SECs, on the left plot.
-20 coefficients were extracted from the spectral envelope obtained following the deconvolution of the CQT spectrogram of the musical signal shown in Figure \ref{fig:deconvolution}.
-The coefficients are shown in dB to be comparable to the envelope which was also shown in dB.
+20 coefficients were extracted from the envelope obtained following the deconvolution of the CQT spectrogram of the musical signal shown in Figure \ref{fig:deconvolution}.
 These coefficients essentially correspond to the harmonics of the musical instrument which contain most of its spectral energy and can therefore be a good representation of its timbre.
 For comparison, we also show the MFCCs computed from the same musical signal, on the right plot.
-20 coefficients were computed using
+20 coefficients were computed using librosa \cite{mcfee2015}, with a sampling rate of 16 kHz, a window length of 1024 samples, and hop length of 512 samples.
+
+We recall below how the MFCCs are commonly derived \cite{mermelstein1976}.
+First, a classic spectrogram is computed from the audio signal using the short-time Fourier transform (STFT) and the powers of the frequencies in each frame are then mapped onto the mel scale \cite{stevens1937} using triangular overlapping windows.
+This essentially produces a mel spectrogram.
+Then, the logs of all the mel frequencies are taken, followed by the discrete cosine transform (DCT) of each frame.
+Finally, the MFCCs are selected as the lowest coefficients in the resulting spectrum, excluding the very first one (DC component).
+This process is supposed to decouple the envelope from the pitch in the time domain, by extracting the slow-varying time components, which most likely correspond to the envelope and which will become the lowest coefficients, from the fast-varying time components, which most likely correspond to the pitch and which will become the higher coefficients.
 
@@
 \end{figure*}
-
-We recall how the MFCCs are computed ...\cite{mermelstein1976}
-mel scale: \cite{stevens1937}
-
-using librosa \cite{mcfee2015}
 

### 2021-07-01 IEEE SP Magazine/07-2021 - Draft/article.tex
Started working on the example section.

--- 2021-06-30 IEEE SP Magazine/06-2021 - Draft/article.tex
@@
 
-We start with the assumption that a log spectrum $X$, in particular, the CQT spectrum, can be represented as the convolution between a pitch-independent spectral envelope $E$ (which mostly contains the timbre information) and an energy-normalized pitch component $P$ (which mostly contains the pitch information), as shown in Equation \ref{eq:assumption}, where $*$ represents the convolution operation.
+We start with the assumption that a log spectrum $S$, in particular, the CQT spectrum, can be represented as the convolution between a pitch-independent spectral envelope $E$ (which mostly contains the timbre information) and an energy-normalized pitch component $P$ (which mostly contains the pitch information), as shown in Equation \ref{eq:assumption}, where $*$ represents the convolution operation.
 \begin{equation}
 \label{eq:assumption}
-X = E * P
+S = E * P
 \end{equation}
@@
 Assuming that pitch and timbre are independent, this implies that the same musical object at different pitches would have a similar envelope but a shifted pitch component (while two different musical objects at the same pitch would have different envelopes but a similar pitch component).
-This is summarized in Equation \ref{eq:observation1}, where $X$, $E$, $P$ and $X'$, $E'$, $P'$ represent the log spectrum, envelope, and pitch component for a musical object and for a pitch-shifted version of the same musical object, respectively.
+This is summarized in Equation \ref{eq:observation1}, where $S$, $E$, $P$ and $S'$, $E'$, $P'$ represent the log spectrum, envelope, and pitch component for a musical object and for a pitch-shifted version of the same musical object, respectively.
 \begin{equation}
@@
 & \begin{cases}
-X = E * P \\
-X' = E' * P' \\
+S = E * P \\
+S' = E' * P' \\
 \end{cases} \\
@@
 
-\emph{Observation 2: the FT of the convolution between two functions is equal to the pointwise product between the FTs of the two functions, a property also known as the convolution theorem \cite{proakis1995}.}
+\emph{Observation 2: the FT of the convolution between two functions is equal to the pointwise product between the FTs of the two functions, a property known as the convolution theorem \cite{proakis1995}.}
 
@@
 & \begin{cases}
-\mathcal{F}(X) = \mathcal{F}(E * P) = \mathcal{F}(E) \cdot \mathcal{F}(P) \\
-\mathcal{F}(X') = \mathcal{F}(E' * P') = \mathcal{F}(E') \cdot \mathcal{F}(P') \\
+\mathcal{F}(S) = \mathcal{F}(E * P) = \mathcal{F}(E) \cdot \mathcal{F}(P) \\
+\mathcal{F}(S') = \mathcal{F}(E' * P') = \mathcal{F}(E') \cdot \mathcal{F}(P') \\
 \end{cases} \\
@@
 & \begin{cases}
-\mathcal{F}(X) = |\mathcal{F}(X)| \cdot e^{j Arg(\mathcal{F}(X))} \\
-\mathcal{F}(X') = |\mathcal{F}(X')| \cdot e^{j Arg(\mathcal{F}(X'))} \\
+\mathcal{F}(S) = |\mathcal{F}(S)| \cdot e^{j Arg(\mathcal{F}(S))} \\
+\mathcal{F}(S') = |\mathcal{F}(S')| \cdot e^{j Arg(\mathcal{F}(S'))} \\
 \end{cases} \\
-& \Rightarrow |\mathcal{F}(X)| \approx |\mathcal{F}(X')|
+& \Rightarrow |\mathcal{F}(S)| \approx |\mathcal{F}(S')|
 \end{split}
@@
 
-Given the previous observations, we can therefore conclude that the FT of the envelope could be approximated by the magnitude of the FT of the log spectrum, while the FT of the pitch component could be approximated by the phase component.
+Given the previous observations, we can therefore conclude that the FT of the envelope could be approximated by the magnitude FT of the log spectrum, while the FT of the pitch component could be approximated by the phase component.
 This finally gives us the estimates for the envelope and the pitch component, after taking their inverse FTs, as shown in Equation \ref{eq:conclusion}, where $\mathcal{F}^{-1}(.)$ represents the inverse FT function.
@@
 \begin{cases}
-\mathcal{F}(E) \approx |\mathcal{F}(X)| \\
-\mathcal{F}(P) \approx e^{j Arg(\mathcal{F}(X))} \\
+\mathcal{F}(E) \approx |\mathcal{F}(S)| \\
+\mathcal{F}(P) \approx e^{j Arg(\mathcal{F}(S))} \\
 \end{cases} \\
@@
 \begin{cases}
-E \approx \mathcal{F}^{-1}(|\mathcal{F}(X)|) \\
-P \approx \mathcal{F}^{-1}(e^{j Arg(\mathcal{F}(X))}) \\
+E \approx \mathcal{F}^{-1}(|\mathcal{F}(S)|) \\
+P \approx \mathcal{F}^{-1}(e^{j Arg(\mathcal{F}(S))}) \\
 \end{cases}
@@
 As we can see, the envelope looks as if the CQT spectrogram has been normalized in pitch, with all the notes being brought down to the lowest frequency (corresponding to C1); while the pitch component looks as if the CQT spectrogram has been stripped down from all its energy, leaving mostly the fundamental frequencies of the notes.
-Note that, in practice, we use a power CQT spectrogram (magnitude to the power of 2) and take the real part of the envelope and pitch component to ensure real values.
-We also note that this deconvolution can potentially be further refined, for example, by zeroing the few negative values in the pitch component and using it to re-derive the spectral envelope from the log spectrum.
+Note that in practice, we use a power CQT spectrogram (i.e., magnitude to the power of 2) and take the real part of the envelope and pitch component to ensure real values.
+We can also potentially post-process this deconvolution, for example, by zeroing the few negative values in the pitch component and using it to derive a refined  envelope from the log spectrum.
 
@@
 
-The spectral envelope resulting from the deconvolution of the CQT spectrum can be thought of as a pitch-normalized CQT spectrum where the spectral content, in particular, the harmonics which represent most of the energy of the musical instrument, has been essentially brought down to the same lowest note level.
+The envelope resulting from the deconvolution of the CQT spectrum can be thought of as a pitch-normalized CQT spectrum where the spectral content, in particular, the harmonics which represent most of the energy of the musical instrument, has been essentially brought down to the same lowest note level.
 Given the octave resolution which was used when computing the CQT spectrum, i.e., the number of bins per octave, we can then easily infer the locations of those harmonics in the envelope \cite{brown1991, brown1992}.
-We can subsequently extract these harmonics or \textit{spectral envelope coefficients} from the envelope and therefore obtain a compact and interpretable feature for characterizing the timbre of the musical instrument.
-Equation \ref{eq:extraction} shows how to derive the indices of the spectral envelope coefficients given $O_r$, the octave resolution, and $N_c$, the number of desired coefficients, and consequently extract the CQT-SECs from the envelope $E$, with $\log_2(.)$ and $round(.)$ representing the binary logarithm and the round function, respectively.
+We can subsequently extract these harmonics or \textit{spectral envelope coefficients} from the envelope and thus obtain a compact and interpretable feature for characterizing the timbre of the musical instrument.
+Equation \ref{eq:extraction} shows how to derive the indices of the spectral envelope coefficients given $O_r$, the octave resolution, and $N_c$, the number of desired coefficients, and finally extract the CQT-SECs from the envelope $E$, with $\log_2(.)$ and $round(.)$ representing the binary logarithm and the round function, respectively.
 \begin{equation}
@@
 Figure \ref{fig:extraction} shows an example of CQT-SECs, on the left plot.
-20 coefficients were extracted from the envelope obtained following the deconvolution of the CQT spectrogram of the musical signal shown in Figure \ref{fig:deconvolution}.
+20 coefficients were extracted from the envelope resulting from the deconvolution of the CQT spectrogram of the musical signal shown in Figure \ref{fig:deconvolution}.
 These coefficients essentially correspond to the harmonics of the musical instrument which contain most of its spectral energy and can therefore be a good representation of its timbre.
@@
 
-\section{Numerical Examples} % or "V - Computational Example"
+\section{A Computational Example}
 
-\subsection{Analysis on an Example}
+We propose to measure the discriminative power of the CQT-SECs by evaluating them on the NSynth dataset, a large-scale of annotated musical notes which is publicly available \cite{engel2017}.
 
-%- The envelope component can additionally be refined, along with the pitch component
-
-\subsection{Comparison on a Dataset}
-
-%Figure \ref{fig:dft_kernels} shows ...
-
-%\begin{figure}[htbp]
-%	\centering
-%	\includegraphics[width=\columnwidth]{dft_kernels.png}
-%	\caption{Kernels derived from the Hanning (top-left), Blackman (top-right), triangular (center-left), Parzen (center-right), Gaussian (with $\alpha = 2.5$) (bottom-left), and Kaiser (with $\beta = 0.5$) (bottom-right) windows.
-The kernels were derived for an $N$-point DFT where $N = 2048$ samples.
-Only the first 100 coefficients at the bottom-left corner of the $N$-by-$N$ kernels are shown.
-The values are displayed in log of amplitude.}
-%	\label{fig:dft_kernels}
-%\end{figure}
+Explain NSynth
+Compare to the MFCCs
+CQT-SECs computed with some parameters
+MFCCs with some parameters
 

### 2021-07-06 IEEE SP Magazine/07-2021 - Draft/article.tex
Working on the computational example section.

@@
 In particular, they are derived through an old process which makes use of the mel scale, a perceptual scale experimentally designed 85 years ago to approximate the human auditory system's response \cite{stevens1937}.
-More recently, a number of data-driven approaches attempted to learn some timbral representations from musical data, but typically in terms of implicit embeddings which are tied to a specific trained model \cite{engel2017}, \cite{pons2017}, and not necessarily as explicit and interpretable features such as the MFCCs, which are still usually preferred as the go-to feature to characterize timbre by MIR practitioners.
+More recently, a number of data-driven approaches attempted to learn some timbral representations from musical data, but typically in terms of implicit embeddings which are tied to a specific trained model \cite{engel2017}, \cite{pons2017}, \cite{hung2019}, \cite{luo2019}, \cite{lee2020} and not necessarily as explicit and interpretable features such as the MFCCs, which are still usually preferred as the go-to feature to characterize timbre by MIR practitioners.
 
@@
 
-We propose the CQT-SECs, a novel timbre feature which is well adapted to musical data, pitch-independent, simple to compute, interpretable, and functional.
+We present the CQT-SECs, a novel timbre feature which is well adapted to musical data, pitch-independent, simple to compute, interpretable, and functional.
 We will show how to derive it from the CQT, a frequency transform with a logarithmic resolution which matches the notes of the equal temperament, a musical scale typically used in Western music \cite{brown1991, brown1992}.
@@
 20 coefficients were extracted from the envelope resulting from the deconvolution of the CQT spectrogram of the musical signal shown in Figure \ref{fig:deconvolution}.
-These coefficients essentially correspond to the harmonics of the musical instrument which contain most of its spectral energy and can therefore be a good representation of its timbre.
+These coefficients essentially correspond to the harmonics of the musical instrument which contain most of its spectral energy and can therefore be a reasonable representation of its timbre.
 For comparison, we also show the MFCCs computed from the same musical signal, on the right plot.
-20 coefficients were computed using librosa \cite{mcfee2015}, with a sampling rate of 16 kHz, a window length of 1024 samples, and hop length of 512 samples.
+20 coefficients were computed using librosa \cite{mcfee2015}, with a sampling rate of 16 kHz, a window length of 1024 samples, and hop length of 512 samples, matching the time resolution of the CQT-SECs so that both features have the same size in time and frequency.
 
-We recall below how the MFCCs are commonly derived \cite{mermelstein1976}.
+We recall below how the MFCCs are commonly derived.
 First, a classic spectrogram is computed from the audio signal using the short-time Fourier transform (STFT) and the powers of the frequencies in each frame are then mapped onto the mel scale \cite{stevens1937} using triangular overlapping windows.
@@
 Finally, the MFCCs are selected as the lowest coefficients in the resulting spectrum, excluding the very first one (DC component).
-This process is supposed to decouple the envelope from the pitch in the time domain, by extracting the slow-varying time components, which most likely correspond to the envelope and which will become the lowest coefficients, from the fast-varying time components, which most likely correspond to the pitch and which will become the higher coefficients.
+This process is supposed to decouple the envelope from the pitch in the time domain, by extracting the slow-varying time components, which most likely correspond to the envelope and which will become the lowest coefficients, from the fast-varying time components, which most likely correspond to the pitch and which will become the highest coefficients \cite{mermelstein1976}.
 
@@
 
-We propose to measure the discriminative power of the CQT-SECs by evaluating them on the NSynth dataset, a large-scale of annotated musical notes which is publicly available \cite{engel2017}.
+We propose to measure the discriminative power of the CQT-SECs by evaluating them on the NSynth dataset, a large-scale and high-quality dataset of annotated musical notes which is publicly available\footnote{https://magenta.tensorflow.org/datasets/nsynth} \cite{engel2017}.
+The NSynth dataset is composed of 305,979 musical notes which were generated from 1,006 instruments as 4-second, monophonic audio signals at a sampling rate of 16 kHz, with pitches ranging over all the note numbers of a standard MIDI piano (21-108) and with 5 different velocities (25, 50, 75, 100, and 127), whenever applicable.
+The instruments are organized into 11 families (bass, brass, flute, guitar, keyboard, mallet, organ, reed, string, synth\_lead, and vocal) and 3 sources (acoustic, electronic, and synthetic).
+We chose to evaluate the CQT-SECs on the notes with a velocity of 75 only, leading to 60,388 different notes for 945 different instruments.
 
-Explain NSynth
-Compare to the MFCCs
-CQT-SECs computed with some parameters
-MFCCs with some parameters
+We derived the CQT-SECs for all the notes in this subset, by first computing the CQT spectrogram using librosa, with a sampling rate of 16 kHz, a hop length of 512 samples, a minimum frequency of 32.70 Hz (corresponding to C1), 95 frequency bins, and 12 bins per octave, and then extracting 20 coefficients from the envelope resulting from the deconvolution of the CQT spectrogram.
+We used a power CQT spectrogram (i.e., magnitude to the power of 2) and took the real part of the envelope to ensure real values.
+For comparison, we also computed the MFCCs using librosa, with a sampling rate of 16 kHz, a window length of 1024 samples, a hop length of 512 samples, and 20 coefficients, matching the time resolution of the CQT-SECs so that both features have the same size in time and frequency.
 

### 2021-07-06 IEEE SP Magazine/07-2021 - Draft/article.tex
Update article.tex

@@
 We used a power CQT spectrogram (i.e., magnitude to the power of 2) and took the real part of the envelope to ensure real values.
-For comparison, we also computed the MFCCs using librosa, with a sampling rate of 16 kHz, a window length of 1024 samples, a hop length of 512 samples, and 20 coefficients, matching the time resolution of the CQT-SECs so that both features have the same size in time and frequency.
+For comparison, we also computed the MFCCs, using librosa, with a sampling rate of 16 kHz, a window length of 1024 samples, a hop length of 512 samples, and 20 coefficients, matching the time resolution of the CQT-SECs so that both features have the same size in time and frequency.
 
@@
 \textit{\textbf{Zafar Rafii}} (zafarrafii@gmail.com) received a PhD in Electrical Engineering and Computer Science from Northwestern University in 2014, and an MS in Electrical Engineering from both Ecole Nationale Superieure de l’Electronique et de ses Applications in France and Illinois Institute of Technology in the US in 2006.
-He is currently a senior research engineer at Gracenote in the US.
+He is currently a research engineer manager at Gracenote in the US.
 He also worked as a research engineer at Audionamix in France.

### 2021-07-13 IEEE SP Magazine/07-2021 - Draft/article.tex
Working on the last section of the article.

@@
 
-We present the CQT-SECs, a novel timbre feature which is well adapted to musical data, pitch-independent, simple to compute, interpretable, and functional.
+We present here the CQT-SECs, a novel timbre feature which is well adapted to musical data, pitch-independent, simple to compute, interpretable, and functional.
 We will show how to derive it from the CQT, a frequency transform with a logarithmic resolution which matches the notes of the equal temperament, a musical scale typically used in Western music \cite{brown1991, brown1992}.
@@
 
-We recall below how the MFCCs are commonly derived.
-First, a classic spectrogram is computed from the audio signal using the short-time Fourier transform (STFT) and the powers of the frequencies in each frame are then mapped onto the mel scale \cite{stevens1937} using triangular overlapping windows.
-This essentially produces a mel spectrogram.
-Then, the logs of all the mel frequencies are taken, followed by the discrete cosine transform (DCT) of each frame.
-Finally, the MFCCs are selected as the lowest coefficients in the resulting spectrum, excluding the very first one (DC component).
-This process is supposed to decouple the envelope from the pitch in the time domain, by extracting the slow-varying time components, which most likely correspond to the envelope and which will become the lowest coefficients, from the fast-varying time components, which most likely correspond to the pitch and which will become the highest coefficients \cite{mermelstein1976}.
-
 \begin{figure*}[htp]
@@
 \end{figure*}
+
+We recall below how the MFCCs are commonly derived.
+First, a classic spectrogram is computed from the audio signal using the short-time Fourier transform (STFT) and the powers of the frequencies in each frame are then mapped onto the mel scale \cite{stevens1937} using triangular overlapping windows.
+This essentially produces a mel spectrogram.
+Then, the log of every mel frequency is taken, followed by the discrete cosine transform (DCT) of each frame.
+Finally, the MFCCs are selected as the lowest coefficients in the resulting spectrum, excluding the very first one (DC component).
+This process is meant to decouple the envelope from the pitch in the time domain, by extracting the slow-varying time components, which most likely correspond to the envelope and which will become the lowest coefficients, from the fast-varying time components, which most likely correspond to the pitch and which will become the highest coefficients \cite{mermelstein1976}.
 
@@
 
-We derived the CQT-SECs for all the notes in this subset, by first computing the CQT spectrogram using librosa, with a sampling rate of 16 kHz, a hop length of 512 samples, a minimum frequency of 32.70 Hz (corresponding to C1), 95 frequency bins, and 12 bins per octave, and then extracting 20 coefficients from the envelope resulting from the deconvolution of the CQT spectrogram.
+We derived the CQT-SECs for all the notes in this subset, by first computing the CQT spectrogram using librosa, with a sampling rate of 16 kHz, a hop length of 512 samples, a minimum frequency of 32.70 Hz (corresponding to C1), 95 frequency bins, and 12 bins per octave, and then extracting 20 coefficients from the envelope resulting from the deconvolution of the CQT spectrogram, leading to CQT-SECs of size 20 coefficients and 126 time frames.
 We used a power CQT spectrogram (i.e., magnitude to the power of 2) and took the real part of the envelope to ensure real values.
-For comparison, we also computed the MFCCs, using librosa, with a sampling rate of 16 kHz, a window length of 1024 samples, a hop length of 512 samples, and 20 coefficients, matching the time resolution of the CQT-SECs so that both features have the same size in time and frequency.
+For comparison, we also computed the MFCCs, using librosa, with a sampling rate of 16 kHz, a window length of 1024 samples, a hop length of 512 samples, and 20 coefficients, leading to MFCCs of size 20 coefficients and 126 time frames, matching the size of the CQT-SECs.
+
+We computed the cosine similarity for every pair of CQT-SECs and every pair of MFCCs (without repetition), after flattening the features into one-dimensional vectors of length 2520 (20 times 126).
+We then averaged these note similarities over every instrument, leading to similarity matrices of size 945 by 945, for both the CQT-SECs and the MFCCs.
+Figure \ref{fig:instrument_similarities} shows the similarity matrices derived for all the pairs of instruments for the CQT-SECs and the MFCCs.
+As we can see, the similarities for the CQT-SECs have more variance, while the similarities for the MFCCs are mostly very high (close to 1) showing poor discriminative power.
+Figure \ref{fig:instrument_similarities2} shows the self-similarities (in green) (i.e., the diagonal of the similarity matrix) and the error bars for the cross-similarities (i.e., means in red and standard deviations in yellow) for every instrument, for the CQT-SECs and the MFCCs.
+As we can see, the self-similarities for the CQT-SECs are noticeably higher that the mean cross-similarities for most of the instruments, and generally higher that the errors too, showing good discriminative power, while the self-similarities and cross-similarities for the MFCCs are all very high (close to 1).
+
+\begin{figure*}[htp]
+    \centering
+    \includegraphics[width=\textwidth]{instrument_similarities.png}
+    \caption{Similarity matrices computed for every pair of notes and averaged over every instrument of the NSynth subset, for the CQT-SECs (left plot) and the MFCCs (right plot).}
+    \label{fig:instrument_similarities}
+\end{figure*}
+
+\begin{figure*}[htp]
+    \centering
+    \includegraphics[width=\textwidth]{instrument_similarities2.png}
+    \caption{Self-similarities and cross-similarities derived from the similarity matrices shown in Figure \ref{fig:instrument_similarities}, for the CQT-SECs (top plot) and the MFCCs (bottom plot).}
+    \label{fig:instrument_similarities2}
+\end{figure*}
+
+We also averaged the note similarities over every instrument family, leading to similarity matrices of size 11 by 11, for both the CQT-SECs and the MFCCs.
+
+\begin{figure*}[htp]
+    \centering
+    \includegraphics[width=\textwidth]{family_similarities.png}
+    \caption{Blah.}
+    \label{fig:instrument_similarities}
+\end{figure*}
+
+\begin{figure*}[htp]
+    \centering
+    \includegraphics[width=\textwidth]{family_similarities2.png}
+    \caption{Blah.}
+    \label{fig:f_similarities2}
+\end{figure*}
 

### 2021-07-14 IEEE SP Magazine/07-2021 - Draft/article.tex
Finishing the computational example section.

@@
 
-We derived the CQT-SECs for all the notes in this subset, by first computing the CQT spectrogram using librosa, with a sampling rate of 16 kHz, a hop length of 512 samples, a minimum frequency of 32.70 Hz (corresponding to C1), 95 frequency bins, and 12 bins per octave, and then extracting 20 coefficients from the envelope resulting from the deconvolution of the CQT spectrogram, leading to CQT-SECs of size 20 coefficients and 126 time frames.
+We derived the CQT-SECs for all the notes in this subset, by first computing the CQT spectrogram using librosa, with a sampling rate of 16 kHz, a hop length of 512 samples, a minimum frequency of 32.70 Hz (corresponding to C1), 95 frequency bins, and 12 bins per octave, and then extracting 20 coefficients from the envelope resulting from the deconvolution of the CQT spectrogram, leading to CQT-SECs of size 20 by 126 (time frames).
 We used a power CQT spectrogram (i.e., magnitude to the power of 2) and took the real part of the envelope to ensure real values.
-For comparison, we also computed the MFCCs, using librosa, with a sampling rate of 16 kHz, a window length of 1024 samples, a hop length of 512 samples, and 20 coefficients, leading to MFCCs of size 20 coefficients and 126 time frames, matching the size of the CQT-SECs.
+For comparison, we also computed the MFCCs, using librosa, with a sampling rate of 16 kHz, a window length of 1024 samples, a hop length of 512 samples, and 20 coefficients, leading to MFCCs of size 20 by 126, just like with the CQT-SECs.
 
-We computed the cosine similarity for every pair of CQT-SECs and every pair of MFCCs (without repetition), after flattening the features into one-dimensional vectors of length 2520 (20 times 126).
-We then averaged these note similarities over every instrument, leading to similarity matrices of size 945 by 945, for both the CQT-SECs and the MFCCs.
-Figure \ref{fig:instrument_similarities} shows the similarity matrices derived for all the pairs of instruments for the CQT-SECs and the MFCCs.
-As we can see, the similarities for the CQT-SECs have more variance, while the similarities for the MFCCs are mostly very high (close to 1) showing poor discriminative power.
-Figure \ref{fig:instrument_similarities2} shows the self-similarities (in green) (i.e., the diagonal of the similarity matrix) and the error bars for the cross-similarities (i.e., means in red and standard deviations in yellow) for every instrument, for the CQT-SECs and the MFCCs.
-As we can see, the self-similarities for the CQT-SECs are noticeably higher that the mean cross-similarities for most of the instruments, and generally higher that the errors too, showing good discriminative power, while the self-similarities and cross-similarities for the MFCCs are all very high (close to 1).
+We computed the cosine similarity for every pair of CQT-SECs and every pair of MFCCs (without repetition), after flattening the features into vectors of length 2520 (20 times 126).
+We then averaged these note similarities over every one of the 945 different instruments, leading to similarity matrices of size 945 by 945, for both the CQT-SECs and the MFCCs.
+Figure \ref{fig:instrument_similarities} shows these instrument similarity matrices for the CQT-SECs and the MFCCs.
+As we can see, the similarities for the CQT-SECs have more variance, while the similarities for the MFCCs are mostly very high (close to 1) showing poor discriminative power between different instruments.
+Figure \ref{fig:instrument_similarities2} additionally shows the self-similarities, i.e., the diagonal in the similarity matrix (in green) and the error bars for the cross-similarities, i.e., the off-diagonal values in the similarity matrix (means in red and standard deviations in yellow) for every instrument, for the CQT-SECs and the MFCCs.
+As we can see, the self-similarities for the CQT-SECs are noticeably higher that the mean of the cross-similarities for most of the instruments, and generally higher that the errors as well, showing good discriminative power, while the self-similarities and cross-similarities for the MFCCs are all very high (close to 1).
 
@@
     \includegraphics[width=\textwidth]{instrument_similarities2.png}
-    \caption{Self-similarities and cross-similarities derived from the similarity matrices shown in Figure \ref{fig:instrument_similarities}, for the CQT-SECs (top plot) and the MFCCs (bottom plot).}
+    \caption{Self-similarities and cross-similarities derived from the instrument similarity matrices shown in Figure \ref{fig:instrument_similarities}, for the CQT-SECs (top plot) and the MFCCs (bottom plot).}
     \label{fig:instrument_similarities2}
@@
 
-We also averaged the note similarities over every instrument family, leading to similarity matrices of size 11 by 11, for both the CQT-SECs and the MFCCs.
+We also averaged the note similarities over every one of the 11 instrument families, leading to similarity matrices of size 11 by 11, for both the CQT-SECs and the MFCCs.
+Figure \ref{fig:family_similarities} shows these instrument family similarity matrices for the CQT-SECs and the MFCCs.
+As we can see here as well, the similarities for the CQT-SECs have more variance, ...
+while the similarities for the MFCCs are mostly very high.
+Figure \ref{fig:family_similarities2} additionally shows the self-similarities (in green) and the error bars for the cross-similarities (means and standard deviations in red) for every instrument family, for the CQT-SECs and the MFCCs.
+As we can see, the self-similarities for the CQT-SECs are noticeably higher that the mean of the cross-similarities for most of the instrument families, and generally higher that the errors as well, while the self-similarities and cross-similarities for the MFCCs are all very high.
 
@@
     \includegraphics[width=\textwidth]{family_similarities.png}
-    \caption{Blah.}
-    \label{fig:instrument_similarities}
+    \caption{Similarity matrices computed for every pair of notes and averaged over every instrument family of the NSynth subset, for the CQT-SECs (left plot) and the MFCCs (right plot).}
+    \label{fig:family_similarities}
 \end{figure*}
@@
     \includegraphics[width=\textwidth]{family_similarities2.png}
-    \caption{Blah.}
-    \label{fig:f_similarities2}
+    \caption{Self-similarities and cross-similarities derived from the instrument family similarity matrices shown in Figure \ref{fig:family_similarities}, for the CQT-SECs (top plot) and the MFCCs (bottom plot).}
+    \label{fig:family_similarities2}
 \end{figure*}

### 2021-07-14 IEEE SP Magazine/07-2021 - Draft/article.tex
Finished the computational example section.

@@
 
-We propose to measure the discriminative power of the CQT-SECs by evaluating them on the NSynth dataset, a large-scale and high-quality dataset of annotated musical notes which is publicly available\footnote{https://magenta.tensorflow.org/datasets/nsynth} \cite{engel2017}.
+We propose to measure the discriminative power of the CQT-SECs by evaluating them for a simple instrument similarity task on the NSynth dataset \cite{engel2017}, a large-scale and high-quality dataset of annotated musical notes which is publicly available\footnote{https://magenta.tensorflow.org/datasets/nsynth}.
 The NSynth dataset is composed of 305,979 musical notes which were generated from 1,006 instruments as 4-second, monophonic audio signals at a sampling rate of 16 kHz, with pitches ranging over all the note numbers of a standard MIDI piano (21-108) and with 5 different velocities (25, 50, 75, 100, and 127), whenever applicable.
-The instruments are organized into 11 families (bass, brass, flute, guitar, keyboard, mallet, organ, reed, string, synth\_lead, and vocal) and 3 sources (acoustic, electronic, and synthetic).
-We chose to evaluate the CQT-SECs on the notes with a velocity of 75 only, leading to 60,388 different notes for 945 different instruments.
+The instruments are organized into 11 families, namely, bass, brass, flute, guitar, keyboard, mallet, organ, reed, string, synth\_lead, and vocal, and 3 sources, namely, acoustic, electronic, and synthetic.
+We chose to evaluate the CQT-SECs on the notes with a velocity of 75 only, leading to a subset of 60,388 different notes for 945 different instruments.
 
@@
 We used a power CQT spectrogram (i.e., magnitude to the power of 2) and took the real part of the envelope to ensure real values.
-For comparison, we also computed the MFCCs, using librosa, with a sampling rate of 16 kHz, a window length of 1024 samples, a hop length of 512 samples, and 20 coefficients, leading to MFCCs of size 20 by 126, just like with the CQT-SECs.
+For comparison, we also computed the MFCCs, using librosa, with a sampling rate of 16 kHz, a window length of 1024 samples, a hop length of 512 samples, and 20 coefficients, leading to MFCCs of same size as the CQT-SECs, i.e., 20 by 126.
 
-We computed the cosine similarity for every pair of CQT-SECs and every pair of MFCCs (without repetition), after flattening the features into vectors of length 2520 (20 times 126).
-We then averaged these note similarities over every one of the 945 different instruments, leading to similarity matrices of size 945 by 945, for both the CQT-SECs and the MFCCs.
+We then computed the cosine similarity for every pair of CQT-SECs and every pair of MFCCs (without repetition), after flattening the features into vectors of length 2520 (20 times 126).
+We averaged these note similarities over every one of the 945 instruments, leading to similarity matrices of size 945 by 945, for both the CQT-SECs and the MFCCs.
 Figure \ref{fig:instrument_similarities} shows these instrument similarity matrices for the CQT-SECs and the MFCCs.
-As we can see, the similarities for the CQT-SECs have more variance, while the similarities for the MFCCs are mostly very high (close to 1) showing poor discriminative power between different instruments.
-Figure \ref{fig:instrument_similarities2} additionally shows the self-similarities, i.e., the diagonal in the similarity matrix (in green) and the error bars for the cross-similarities, i.e., the off-diagonal values in the similarity matrix (means in red and standard deviations in yellow) for every instrument, for the CQT-SECs and the MFCCs.
-As we can see, the self-similarities for the CQT-SECs are noticeably higher that the mean of the cross-similarities for most of the instruments, and generally higher that the errors as well, showing good discriminative power, while the self-similarities and cross-similarities for the MFCCs are all very high (close to 1).
+As we can see, the similarities for the CQT-SECs have more variance, while the similarities for the MFCCs are mostly very high (close to 1) showing poor discriminative power between the different instruments.
+Figure \ref{fig:instrument_similarities2} additionally shows the self-similarities, i.e., the diagonal values in the similarity matrix (in green) and the error bars for the cross-similarities, i.e., the off-diagonal values in the similarity matrix (means in red and standard deviations in yellow) for every instrument, for the CQT-SECs and the MFCCs.
+As we can see, the self-similarities for the CQT-SECs are noticeably higher than the means of the cross-similarities for most of the instruments, and generally higher that the means plus standard deviations as well, showing good discriminative power, while the self-similarities and cross-similarities for the MFCCs are all very high (close to 1).
 
@@
 Figure \ref{fig:family_similarities} shows these instrument family similarity matrices for the CQT-SECs and the MFCCs.
-As we can see here as well, the similarities for the CQT-SECs have more variance, ...
-while the similarities for the MFCCs are mostly very high.
+As we can see here as well, the similarities for the CQT-SECs have more variance, showing mostly clearer high self-similarities and low cross-similarities, while the similarities for the MFCCs are mostly very high.
 Figure \ref{fig:family_similarities2} additionally shows the self-similarities (in green) and the error bars for the cross-similarities (means and standard deviations in red) for every instrument family, for the CQT-SECs and the MFCCs.
-As we can see, the self-similarities for the CQT-SECs are noticeably higher that the mean of the cross-similarities for most of the instrument families, and generally higher that the errors as well, while the self-similarities and cross-similarities for the MFCCs are all very high.
+As we can again see, the self-similarities for the CQT-SECs are noticeably and mostly higher than the means plus standard deviations (except for the string and synth\_lead families), while the self-similarities and cross-similarities for the MFCCs are all very high and harder to tell apart.
 

### 2021-07-14 IEEE SP Magazine/07-2021 - Draft/article.tex
Finished article!

@@
 It is typically associated with the spectral (but also the temporal) envelope and assumed to be independent from the pitch (but also the loudness) of the sound \cite{moore2004}.
-In this article, we will show how to design a simple but functional pitch-independent timbre feature which is well adapted to musical data, by deriving it from the constant-Q transform (CQT), a log-frequency transform which matches the equal tempered musical scale \cite{brown1991, brown1992}.
+In this article, we will show how to design a simple but functional pitch-independent timbre feature which is well adapted to musical data, by deriving it from the constant-Q transform (CQT), a log-frequency transform which matches the equal-tempered musical scale \cite{brown1991, brown1992}.
 We will show how to decompose the CQT spectrum into an energy-normalized pitch component and a pitch-independent spectral envelope, the latter from which we will extract a number of timbral coefficients.
@@
 Additionally, the ability to decompose the spectrum of a sound (here, the CQT spectrum) into a pitch-independent spectral envelope and an energy-normalized pitch component can be useful for analysis, transformation, and resynthesis of music signals.
-The energy-normalized pitch component can also potentially be used for tasks such as pitch identification, melody extraction, and chord recognition.
+The energy-normalized pitch component can also potentially be used for tasks such as pitch identification and melody extraction.
 
@@
 In particular, they are derived through an old process which makes use of the mel scale, a perceptual scale experimentally designed 85 years ago to approximate the human auditory system's response \cite{stevens1937}.
-More recently, a number of data-driven approaches attempted to learn some timbral representations from musical data, but typically in terms of implicit embeddings which are tied to a specific trained model \cite{engel2017}, \cite{pons2017}, \cite{hung2019}, \cite{luo2019}, \cite{lee2020} and not necessarily as explicit and interpretable features such as the MFCCs, which are still usually preferred as the go-to feature to characterize timbre by MIR practitioners.
+More recently, a number of data-driven approaches attempted to learn some timbral representations from musical data, but typically in terms of implicit embeddings which are tied to specific trained models \cite{engel2017, pons2017, hung2019, luo2019, lee2020} and not necessarily as explicit and interpretable features such as the MFCCs, which are still usually preferred as the go-to feature to characterize timbre by MIR practitioners.
 
@@
 
-We present here the CQT-SECs, a novel timbre feature which is well adapted to musical data, pitch-independent, simple to compute, interpretable, and functional.
-We will show how to derive it from the CQT, a frequency transform with a logarithmic resolution which matches the notes of the equal temperament, a musical scale typically used in Western music \cite{brown1991, brown1992}.
+We hereby present the CQT-SECs, a novel timbre feature which is well adapted to musical data, pitch-independent, simple to compute, interpretable, and functional.
+We will show how to derive it from the CQT, a frequency transform with a logarithmic resolution which matches the notes of the equal temperament, a tuning system typically used in Western music \cite{brown1991, brown1992}.
 We will first show how to decompose the CQT spectrum into a pitch-independent spectral envelope and an energy-normalized pitch component, and then extract a number of timbral coefficients from the spectral envelope.
@@
 \end{equation}
-
 This convolution process can be thought of as a source-filter model \cite{fant1970} which is here not applied in the time domain but in the frequency domain, with the source and the filter being the pitch component and the envelope, respectively.
@@
 Figure \ref{fig:deconvolution} shows an example of deconvolution of a CQT spectrogram into its envelope and pitch component.
-The CQT spectogram was computed from an audio signal created by concatenating 12 4-second notes of an acoustic bass playing from C1 (32.70 Hz) to B1 (61.74 Hz) in ascending order.
+The CQT spectogram was computed from an audio signal created by concatenating 12 4-second notes of an acoustic bass playing from C1 (32.70 Hz) to B1 (61.74 Hz) in ascending pitch.
 The notes come from the NSynth dataset \cite{engel2017} and correspond to instrument id \texttt{bass\_acoustic\_000}, MIDI numbers \texttt{024} to \texttt{035}, and velocity number \texttt{075}.
-The CQT spectrogram was computed using librosa \cite{mcfee2015,schoerkhuber2010}, with a sampling rate of 16 kHz, a hop length of 512 samples, a minimum frequency of 32.70 Hz (corresponding to C1), 95 frequency bins, and 12 bins per octave.
-As we can see, the envelope looks as if the CQT spectrogram has been normalized in pitch, with all the notes being brought down to the lowest frequency (corresponding to C1); while the pitch component looks as if the CQT spectrogram has been stripped down from all its energy, leaving mostly the fundamental frequencies of the notes.
-Note that in practice, we use a power CQT spectrogram (i.e., magnitude to the power of 2) and take the real part of the envelope and pitch component to ensure real values.
+The CQT spectrogram was computed using librosa \cite{mcfee2015, schoerkhuber2010}, with a sampling rate of 16 kHz, a hop length of 512 samples, a minimum frequency of 32.70 Hz (corresponding to C1), 95 frequency bins, and 12 bins per octave.
+As we can see, the envelope looks as if the CQT spectrogram has been normalized in pitch, with all the notes being brought down to the lowest frequency (corresponding to C1); while the pitch component looks as if the CQT spectrogram has been stripped down from its energy, leaving mostly the fundamental frequencies of the notes.
+Note that in practice, we use a power CQT spectrogram (i.e., magnitude to the power of 2) and take the real part for the envelope and pitch component to ensure real values.
 We can also potentially post-process this deconvolution, for example, by zeroing the few negative values in the pitch component and using it to derive a refined  envelope from the log spectrum.
@@
 
-Figure \ref{fig:extraction} shows an example of CQT-SECs, on the left plot.
+Figure \ref{fig:extraction} shows an example of CQT-SECs (on the left plot).
 20 coefficients were extracted from the envelope resulting from the deconvolution of the CQT spectrogram of the musical signal shown in Figure \ref{fig:deconvolution}.
 These coefficients essentially correspond to the harmonics of the musical instrument which contain most of its spectral energy and can therefore be a reasonable representation of its timbre.
-For comparison, we also show the MFCCs computed from the same musical signal, on the right plot.
+For comparison, we also show the MFCCs computed from the same musical signal (on the right plot).
 20 coefficients were computed using librosa \cite{mcfee2015}, with a sampling rate of 16 kHz, a window length of 1024 samples, and hop length of 512 samples, matching the time resolution of the CQT-SECs so that both features have the same size in time and frequency.
@@
     \includegraphics[width=\textwidth]{extraction.png}
-    \caption{CQT-SECs extracted from the envelope obtained following the deconvolution shown in Figure \ref{fig:deconvolution} (left plot, shown in dB) and MFCCs computed from the same musical signal (right plot).}
+    \caption{CQT-SECs extracted from the envelope obtained following the deconvolution of the CQT spectrogram shown in Figure \ref{fig:deconvolution} (left plot, shown in dB) and MFCCs computed from the same musical signal (right plot).}
     \label{fig:extraction}
@@
 
-We derived the CQT-SECs for all the notes in this subset, by first computing the CQT spectrogram using librosa, with a sampling rate of 16 kHz, a hop length of 512 samples, a minimum frequency of 32.70 Hz (corresponding to C1), 95 frequency bins, and 12 bins per octave, and then extracting 20 coefficients from the envelope resulting from the deconvolution of the CQT spectrogram, leading to CQT-SECs of size 20 by 126 (time frames).
-We used a power CQT spectrogram (i.e., magnitude to the power of 2) and took the real part of the envelope to ensure real values.
+We derived the CQT-SECs for all the notes in this subset, by first computing the CQT spectrogram using librosa, with a sampling rate of 16 kHz, a hop length of 512 samples, a minimum frequency of 32.70 Hz (corresponding to C1), 95 frequency bins, and 12 bins per octave, and then extracting 20 coefficients from the envelope resulting from the deconvolution of the CQT spectrogram, leading to CQT-SECs of size 20 (coefficients) by 126 (time frames).
+We used a power CQT spectrogram (i.e., magnitude to the power of 2) and took the real part for the envelope to ensure real values.
 For comparison, we also computed the MFCCs, using librosa, with a sampling rate of 16 kHz, a window length of 1024 samples, a hop length of 512 samples, and 20 coefficients, leading to MFCCs of same size as the CQT-SECs, i.e., 20 by 126.
@@
 We also averaged the note similarities over every one of the 11 instrument families, leading to similarity matrices of size 11 by 11, for both the CQT-SECs and the MFCCs.
-Figure \ref{fig:family_similarities} shows these instrument family similarity matrices for the CQT-SECs and the MFCCs.
-As we can see here as well, the similarities for the CQT-SECs have more variance, showing mostly clearer high self-similarities and low cross-similarities, while the similarities for the MFCCs are mostly very high.
-Figure \ref{fig:family_similarities2} additionally shows the self-similarities (in green) and the error bars for the cross-similarities (means and standard deviations in red) for every instrument family, for the CQT-SECs and the MFCCs.
-As we can again see, the self-similarities for the CQT-SECs are noticeably and mostly higher than the means plus standard deviations (except for the string and synth\_lead families), while the self-similarities and cross-similarities for the MFCCs are all very high and harder to tell apart.
+Figure \ref{fig:family_similarities} shows these instrument family similarity matrices and figure \ref{fig:family_similarities2} additionally shows the self-similarities (in green) and the cross-similarities (means and standard deviations in red) for every instrument family, for the CQT-SECs and the MFCCs.
+As we can again see, the similarities for the CQT-SECs have more variance compared with the MFCCs, and their self-similarities are noticeably and mostly higher than the means plus standard deviations of their cross-similarities while the self-similarities and cross-similarities for the MFCCs are all very high and harder to tell apart.
+We can clearly see with the CQT-SECs, for example, that the organ family has high self-similarity, showing that most of the instruments and notes in that family are quite similar to each other in terms of timbre.
+We can also see that the organ family is somewhat similar to the brass, flute, and vocal families but quite different from the other families.
+On the other hand, the string family has fairly low self-similarity, showing that either the instruments or notes in that family differ substantially from each other, or the CQT-SECs were not able to properly capture the overall timbre for this family.
 
@@
 
-We have shown that we can derive a simple but function timbre feature which is more adapted to musical data ...
-
-%Limitations:
-%- just like MFCCs, not very robust to noise, to mixture of musical instruments playing at different pitches
+We have shown that we can design a simple but functional pitch-independent timbre feature which is well-adapted to musical data by decomposing the CQT spectrum into a pitch-independent spectral envelope and an energy-normalized pitch component, and extracting a number of coefficients from the spectral envelope.
+These CQT-SECs provide a compact and interpretable feature for characterizing the timbre of a musical instrument, showing higher variance in terms of similarities on a simple instrument similarity task, compared with the MFCCs which are commonly used by MIR practitioners to characterize timbre in music.
+For the interested reader, we provide a Python implementation of the CQT-SEC online with some examples\footnote{https://github.com/zafarrafii/CQT-SEC-Python}.
 
